### 2017-10-11 paper01_peloquin_frank2016/model_walkthru/paper01_peloquin_frank2016.rmd
clean up pf16 notes a tad and put in model walkthru dir

---
title: "paper01 model walkthru"
subtitle: "Peloquin & Frank (2016), 'Determining the alternatives for scalar implicature'"
author: "timothy leffel"
date: "oct20/2017"
---

### Gameplan

- recap of paper 
- what the model is (broad strokes)
- what the model does 
- ...

### Recap of Paper

### Model

```{r}
lefftpack::lazy_setup()
# devtools::install_github("benpeloquin7/rrrsa")
library("rrrsa")

### CONTENTS -------

# SIX DATASETS

# THREE HELPER FUNCTIONS

# SIX ANALYSIS FUNCTIONS

### DATASETS ------------------------------------------------------------------

# note: on any dataset `df`, start by calling `?df` to see info

# example docs for `rrrsa::peloquinFrank_2Alts`
# 
#   This data set contains empirical literal listener semantics and pragmatic
#   judgments for entailment items (two alternatives) from Peloquin & Frank 2016
# 
#   1) quantityVarName :: "stars"
#   2) semanticsVarName :: "speaker.p"
#   3) itemVarName :: "words"
#   4) groupVarName :: "scale"
#   other) experiment number :: "exp"
#   other) Pragmatic judgments study 1 (e6), study 2 (e11)
# 
#   see also: https://github.com/langcog/scalar_implicature

# 136x16 df, data from frank et al. (under review)
rrrsa::d_pragmods

# 50x7 df, data from pel+frk16, 
rrrsa::peloquinFrank_2Alts

# 75x7 df
rrrsa::peloquinFrank_3Alts

# 100x7 df
rrrsa::peloquinFrank_4Alts

# 125x7 df 
rrrsa::peloquinFrank_5Alts

### HELPER FUNCTIONS ----------------------------------------------------------

# converts vec2 to whatever type vec1 is
rrrsa::rsa.convertVecType(vec1, vec2)

# this is just `function(x) x/sum(x)` for positive `x`
rrrsa::rsa.normVec(v)

# just changes desired column names 
rrrsa::rsa.renameCol(df, currNames, replacements)

### ANALYSIS FUNCTIONS --------------------------------------------------------

### `rrrsa::rsa.informativity(m_u, alpha, cost)` -------
# ifelse(m_u == 0, 0, exp(-alpha * (-log(m_u) - cost)))
rrrsa::rsa.informativity(m_u, alpha, cost)

### `rrrsa::rsa.utility(items, costs, alpha)` ------
# rsa.normVec(mapply(rsa.informativity, items, costs, alpha = alpha))
rrrsa::rsa.utility(items, costs, alpha)

### `rrrsa::rsa.reason(m, costs, priors, depth, alpha)` ----
rrrsa::rsa.reason(m, costs, priors, depth, alpha, usePriorEveryRecurse)

### `rrrsa::rsa.runDf()` ----------------
rrrsa::rsa.runDf(data, quantityVarName, semanticsVarName, 
                 itemVarName, costsVarName, priorsVarName, depth, alpha,
                 usePriorEveryRecurse)

### `rrrsa::rsa.tuneDepthAlpha()` --------
rrrsa::rsa.tuneDepthAlpha()

### rrrsa::rsa.fullRecursion(m, costs, priors, alpha) ------
# `m` a matrix
# `costs` a vector of costs, w length == ncol(m) (defaults to zeros)
# `priors` vector of priors, w length == nrow(m) (defaults to ones)
# `alpha` scalar noise param, defaults to 1 
# returns a posterior distribution 
rrrsa::rsa.fullRecursion(m, costs, priors, alpha)

# pdist=runif; nvals=100; ncols=4; params=c(0,1)
fullrec <- function(pdist, nvals, ncols, params){
  set.seed(3369)
  m <- matrix(pdist(nvals, params[1], params[2]), ncol=ncols)
  if (min(m) < 0) message("converting negative vals to positive")
  m <- abs(m); m <- m/sum(m)
  m_post <- rrrsa::rsa.fullRecursion(m=m)
  par(mfrow=c(1,3)); print(hist(m)); print(hist(m_post)); print(plot(m, m_post))
  title(paste0("with nvals=",nvals," ncols=",ncols, 
               "\nparams=",paste(params, collapse=",")))
}
lapply(c(2,4,5,10,100), function(ncols){
  fullrec(pdist=runif, nvals=1e3, ncols=ncols, params=c(0,1))
})
lapply(c(2,4,5,10,100), function(ncols){
  fullrec(pdist=rnorm, nvals=1e3, ncols=ncols, params=c(0,1))
})

```

### 2017-10-11 paper01_peloquin_frank2016/paper_and_materials/rrrsa_vignette.Rmd
clean up pf16 notes a tad and put in model walkthru dir

---
title: "rrrsa"
author: "Ben Peloquin"
date: "[created sept16/2016, compiled by tim, oct06/2017]"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{summary}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Rational speech act model (RSA) of pragmatic inference

`rrrsa` is an R package for running RSA models -- Bayesian models of pragmatic inference. `rrrsa` was created by Ben Peloquin in collaboration with Michael C. Frank and has been optimized for analysis of experimental data such as those presented in Frank, et al. (Under Review) and Peloquin & Frank (2016). For other, more flexible variants of RSA models, please see http://forestdb.org/models/scalar-implicature.html.

## Installation
You can install the latest version of `rrrsa` by installing `devtools` and running:
```{r install_demo, eval=FALSE}
# install.packages("devtools")
# devtools::install_github("benpeloquin7/rrrsa")
```

## What is RSA?
Rational speech act (RSA) models frame language understanding as a special case of social cognition in which `speakers` and `listeners` reason about one another recursively. A `pragmatic listener` $P_{L_n}(m|u)$, reasons about intended meaning $m$ of an utterance $u$ by a `rational speaker` $P_{s_n}(u|m)$ who chooses an utterance according to the expected utility of an utterance $U(m;u)$. $\alpha$ is a decision noise parameter.

$$P_{L_n}(m|u) \propto P_{S_n}(u|m)P(m)$$
$$P_{S_n} \propto e^{U(m;u)}$$
$$U(m;u) = -\alpha(-\log(P_{L_{n-1}}(m|u)) - C(u))$$

See Frank & Goodman (2012) and Goodman & Stuhmuller (2013) for the original descriptions of the RSA framework. Frank, et al. (Under Review) also provides a comprehensive presentation and evaluation of RSA.

## rrrsa includes empirical data
Data from "Rational speech act models of pragmatic reasoning in reference games" (Frank, et al., Under Review) and "Determining the alternatives in scalar implicature" (Peloquin & Frank, 2016) are also included in this package. Examples using data from these studies are included below.

## rrrsa includes access to all model components
`rrrsa` provides users with access to all model components. The following sections demonstrate how this functionality can be used.

```{r libraries, warning=FALSE, message=FALSE}
library(knitr)
library(ggplot2)
library(tidyr)
library(dplyr)
library(purrr)
library(rrrsa)
```

### Calculating the informativity of an utterance with `rsa.informativity()`
`rsa.informativity()` takes three arguments, literal semantics $P_{L_0}$, `alpha` level (default 1), and `cost` (default 0). This function returns the surprisal of an utterance minus cost, multiplied by alpha.
```{r rsa.informativity_demo}
rsa.informativity(0.4)
rsa.informativity(rsa.informativity(0.4), alpha = 2, cost = 0.5)
```

### Calculating the utility of an utterance with `rsa.utility()`
`rsa.utility` takes an input vector of literal listener semantics and outputs a normalized vector of speaker likelihoods. If `costs` are not specified the default 0's vector is used. If `alpha` is not specified a default value of $1$ is used.
```{r rsa.utility_demo}
literalSemantics <- c(0.0, 0.0, 0.3, 0.3, 0.4)
costs <- c(0.0, 0.0, 0.2, 0.3, 0.4)
rsa.utility(items = literalSemantics, costs = costs, alpha = 3)
```

### Computing one full recursion with `rsa.fullRecursion()`
In the RSA framework one full recursion consists of a `pragmatic listner` $P_{L_1}$ who reasons about a `rational speaker` $P_{s_1}$ who reason about a `literal listener` $P_{L_0}$. Expected input is an $m$ matrix of  $P_{L_0}$ `literal listener` values in which columns correspond to items (words) and rows correspond to semantic quantity (i.e. "stars" in Peloquin & Frank, 2016). Optional arguments include a `costs` vector which must be the same length as `ncol` and an optional `priors` vector which must be the same length as `nrows`. `rsa.fullRecursion()` provides safety checking for these cases. Output corresponds with pragmatic listener posterior predictions.
```{r rsa.full_recursion_demo, fig.height=5, fig.width=6}
m <- matrix(data = c(1.0, 0.0, 0.0, 0.0, 0.0,
                     0.0, 0.25, 0.25, 0.25, 0.25,
                     0.0, 0.0, 0.0, 0.0, 1.0), nrow = 5)
colnames(m) <- c("none", "some", "all")
rownames(m) <- 1:5
# costs <- c("none" = 0, "some" = 0, "all" = 0)
# priors <- rnorm(n = nrow(m), mean = 0.5, sd = 0.1)
res <- rsa.fullRecursion(m = m)
res <- as.data.frame(res) %>%
  mutate(quantity = rownames(.))

## Prep data
pragmaticsTidied <- res %>%
  gather(word, pragmatics, -quantity)
semanticsTidied <- as.data.frame(m) %>%
  mutate(quantity = rownames(.)) %>%
  gather(word, semantics, c(none, some, all))
fullData <- merge(pragmaticsTidied, semanticsTidied) %>%
  gather(type, value, c(pragmatics, semantics)) %>%
  mutate(quantity = as.numeric(quantity),
         lineSize = ifelse(word == "some", 2, 1))

## Visualize implicature
ggplot(fullData, aes(x = quantity, y = value, col = word)) +
  geom_line(aes(size=lineSize)) +
  facet_wrap(~type) +
  theme_bw()
```

### Running multiple recursions with `rsa.reason()`
`rsa.reason()` is a wrapper function for `rsa.fullRecursion` which provides an additional `depth` parameter specifying the recursive depth during reasoning. If depth is not provided, default value is $1$.
```{r rsa.reason_demo, results='asis'}
all(rsa.reason(m = m, depth = 2) == rsa.fullRecursion(rsa.fullRecursion(m = m)))
rsa.reason(m = m, depth = 2) %>%
  kable()
```

### Running data frames with `rsa.runDf()`
Run RSA on a tidied data frame and avoid running individual model components individually with `rsa.runDf()`. This is the primary workhorse function of `rrrsa`. An rrrsa-ready, tidied data frame must contain columns for semantic `quantity`, `item` and `semantics`, where each row corresponds with unique `item`/`quantity` combination. A user should specify their naming convention for these items in the `quantityVarName`, `itemVarName` and `semanticsVarName` arguments. The `costVarName` and `priorsVarName` arguments correspond with `costs` and/or `priors` data. Users can specify values for `alpha` and `depth` hyperparamenters. `rsa.runDf()` will return a data frame with model predictions '`preds`' appended as a new column.

```{r rsa.runDf_demo}
## Hypothetical literal listener data we might want to use RSA to simulate implicature.
df <- data.frame(scales = rep("some_all", 15),
                  stars = as.factor(rep(1:5, 3)),
                  starsChar = as.factor(rep(c("one", "two", "three", "four", "five"), 3)),
                  words = c(rep("all", 5), rep("some", 5), rep("none", 5)),
                  listenerSemantics = c(rep(0.0, 4), 1.0,
                                0.0, rep(0.25, 4),
                                1.0, rep(0.0, 4)))
rsa.runDf(df,
          quantityVarName="stars",
          semanticsVarName="listenerSemantics",
          itemVarName="words") %>%
  kable()
```

Importantly, `rsa.runDf()` maintains all column naming and can handle multiple data types. For example, we can run `rsa.runDf()` with a character vector for `quantity` (contrast with the factor vector used above):
```{r rsa.runDf_demo2}
all(rsa.runDf(df, quantityVarName = "starsChar",
              semanticsVarName = "listenerSemantics",
              itemVarName = "words") ==
      rsa.runDf(df, quantityVarName = "stars",
                 semanticsVarName = "listenerSemantics",
                 itemVarName = "words"))
```

A frequent use case for RSA will require running RSA over multiple groups of data. Rather than subsetting data frames and running RSA iteratively, we recommend using `map_df()` from the `purrr` package. Here we split by the `scales` variable.
```{r rsa.runDf_map_df, message=FALSE}
## RSA of literal semantics we'd like to run RSA on
df <- data.frame(scales = c(rep("some_all", 10), rep("good_excellent", 10)),
                  stars = as.factor(rep(1:5, 4)),
                  words = c(rep("all", 5), rep("some", 5), c(rep("excellent", 5), rep("good", 5))),
                  listenerSemantics = c(rep(0.0, 4), 1.0,
                                0.0, rep(0.25, 4),
                                rep(0.0, 4), 1.0,
                                0.0, rep(0.25, 4))) %>% mutate(priors = 0.20)
df$costs <- c(rep(3, 5), rep(4, 5), rep(9, 5), rep(4, 5))

## Using purrr::map_df() run rsa.runDf() over df subsets
df %>%
  split(.$scales) %>%
  map_df(~rsa.runDf(
    data=.x,
    quantityVarName="stars",
    semanticsVarName="listenerSemantics",
    itemVarName="words",
    costsVarName="costs",
    depth=2)) %>%
  head(n=10) %>%
  kable()
```

## An example tuning hyperparameters using `rsa.runDf()` and `purrr::map_df()` with empirical data from Peloquin & Frank (2016)

### Data description for Peloquin & Frank (2016) "Determining the alternatives for scalar implicature"
`rrrsa` includes empirical literal listener $P_{L_0}$ which can be used as input to `rrrsa` as well as $P_{L_1}$ pragmatic judgments for model tuning and comparison. Four data sets are included:

`peloquinFrank_2Alts`: data set with entailment alternatives

`peloquinFrank_3Alts`: data set with entailment alternatives + universal none

`peloquinFrank_4Alts`: data set with entailment alternatives + top two empirically derived alts

`peloquinFrank_5Alts`: data set with entailment alternatives + top two empirically derived alts + neutral valence alternative

```{r peloquinFrank_2Alts_str}
str(peloquinFrank_2Alts) %>%
  kable()
```

See `?peloquinFrank_2alts` for data descriptions and [link to CogSci paper here...]

RSA hyperparameters `alpha` and `depth` can be tuned using custom functions. Here we give a simple example of hyperparamter tuning with grid search using nested for loops. We examine empirical data from Peloquin & Frank (2016).

In this case our grouping variable is `scales` -- we run `rsa.runDf()` on each scale subset for each alpha and depth we're interested in saving the results to `res`.
```{r tuning_hyperparams_demo, warnings=FALSE, messages=FALSE}
## hyperparams
depths <- seq(0, 4)
alphas <- seq(0, 5, by=0.2)

## Grid search using nested for-loops
res <- c()
for (alpha in alphas) {
  for (depth in depths) {
    curr_out <- peloquinFrank_5Alts %>%
      split(.$scale) %>%
      map_df(~rsa.runDf(
        data=.x,
          quantityVarName="stars",
          semanticsVarName="speaker.p",
          itemVarName="words",
          depth=depth,
          alpha=alpha)) %>%
      mutate(alpha=alpha,
             depth=depth)
    res <- rbind(res, curr_out)
  }
}
```

In this case, we're only interested in how the model predicts some of the items (entailment items; see Peloquin & Frank, 2016).
```{r tuning_hyperparams_demo2}
target_items <- c("good", "excellent",
                  "liked", "loved",
                  "memorable", "unforgettable",
                  "palatable", "delicious",
                  "some", "all")

d_cors <- res %>%
  filter(words %in% target_items) %>%
  group_by(alpha, depth) %>%
  summarise(r=cor(e11, preds)) %>%
  ungroup
```

How does the model respond to the hyperparameters?
```{r tuning_hyperparams_plot, fig.height=5, fig.width=6}
ggplot(d_cors, aes(x=alpha, y=r, col=as.factor(depth))) +
  geom_line(alpha=0.8) +
  theme_bw()
```

```{r tuning_hyperparams_demo3, results='asis'}
d_cors %>%
  summarise(max_r=max(r),
            max_alpha=alpha[which.max(r)],
            max_depth=depth[which.max(r)]) %>%
  kable()
```

## Simulating pragmatic inference with data from "Rational speech act models of pragmatic reasoning in reference games" - Frank, et al. (Under Review)

### Data description for Frank, et al. (Under Review)

`rrrsa` includes empirical data used in Frank et al. (Under Review) model simulations in the `rrrsa::d_pragmods` data frame. We provide users with access to the data and include a short example of running simulations.

```{r pragmod_head}
head(d_pragmods) %>%
  kable()
```

You'll notice some `NA` values. This is because empirical data were not measured for all items, however we can supply the model with literal semantics for those items. These are present in the `speaker.p` column.

For more detailed information on the data included in `d_pragmods` run.
```{r d_pragmods_info, eval=FALSE}
?d_pragmods
```

### RSA simulations

Unlike in the previous example, we'd like to run rsa via `rsa.runDf()` over individual experiments rather than scales. We've provided a grouping variable in the data frame (`grouper`) which allows us to do this.

We split by the grouping variable `grouper` and use `purrr::map_df()` which allows us to run `rsa.runDf()` over subsets of the data frame. This equivalent to subsetting `d_pragmods` by each factor in `grouper`, running `rsa.runDf` and `rbind()`ing the results.
```{r pragmods_sim_1}
d_preds_priors <- d_pragmods %>%
  split(list(.$grouper)) %>%
  map_df(~rsa.runDf(
    data=.x,
    quantityVarName="object",
    semanticsVarName="speaker.p",
    itemVarName="query",
    priorsVarName = "priorValue",
    depth=1,
    alpha=1)) %>%
  mutate(keep_indices = ifelse(!is.na(count), "keep", "throwout")) %>%
  filter(keep_indices=="keep")
```
Notice that we're not interested in all the data here. We only collected empirical judgments for some of the items, but `rrrsa` requires that we include all the alternatives. We handle this by identifiying all the rows that contain `NA`s after running the simulations and removing them.

How does our model do?
```{r pragmods_sim_1_plot, fig.height=5, fig.width=6}
ggplot(d_preds_priors, aes(x=preds, y=p, col=expt, pch=object)) +
  geom_pointrange(aes(ymin = cil, ymax = cih)) + 
  xlim(c(0,1)) + ylim(c(0,1)) + 
  ylab("Proportion choosing target") + 
  xlab("Model Predictions") + 
  geom_smooth(method="lm", aes(group=1)) + 
  geom_abline(slope = 1, intercept = 0, lty=2) +
  ggtitle(paste0("Priors model sims\nalpha=1, depth=1\n",
                 "cor=",
                 round(cor(d_preds_priors$p, d_preds_priors$preds), 2))) +
  theme_bw()
```

This data corresponds with the second row of Table 2 in Frank, et al. (Under Review).

We can re-run without including priors by simply omitting the `priors` column name, in this case `priorValue`.

```{r pragmods_sim_2}
d_preds_no_priors <- d_pragmods %>%
  split(list(.$grouper)) %>%
  map_df(~rsa.runDf(
    data=.x,
    quantityVarName="object",
    semanticsVarName="speaker.p",
    itemVarName="query",
    # priorsVarName = "priorValue",
    depth=1,
    alpha=1)) %>%
  mutate(keep_indices = ifelse(!is.na(count), "keep", "throwout")) %>%
  filter(keep_indices=="keep")
```

```{r pragmods_sim_2_plot, fig.height=5, fig.width=6}
ggplot(d_preds_no_priors, aes(x=preds, y=p, col=expt, pch=object)) +
  geom_pointrange(aes(ymin = cil, ymax = cih)) + 
  xlim(c(0,1)) + ylim(c(0,1)) + 
  ylab("Proportion choosing target") + 
  xlab("Model Predictions") + 
  geom_smooth(method="lm", aes(group=1)) + 
  geom_abline(slope = 1, intercept = 0, lty=2) +
  ggtitle(paste0("No priors model sims\nalpha=1, depth=1\n",
                 "cor=",
                 round(cor(d_preds_no_priors$p, d_preds_no_priors$preds), 2))) +
  theme_bw()
```

This data corresponds with the second row of Table 4 in Frank, et al. (Under Review).

### 2017-10-11 paper01_peloquin_frank2016/model_walkthru/notes.r

@@ -1,0 +1,194 @@
+### ---------------------------------------------------------------------------
+# this is a scratchpad for playing with ben peloquin's `rrrsa::` package. 
+# the package contains data/code for peloquin + frank 2016. 
+# 
+# this file can be used as starter code, but a more extensive, well-organized, 
+# careful, etc. walkthru will be made available in advance of the oct20 meeting
+# 
+#  
+# NOTE: to use the `rrrsa::` package, you need to first install it. 
+#       because it is not a fully vetted package (those usually live on CRAN), 
+#       you will need to install it with the `install_github()` function from 
+#       the `devtools::` package. 
+#       aaaand as you might expect, to use that, you gotta install it first. 
+# 
+# 
+#   - tim leffel, oct11/2017
+### ---------------------------------------------------------------------------
+
+
+
+
+### SETUP ---------------------------------------------------------------------
+
+# if you don't know whether you have the `devtools::` package, then first say: 
+# install.packages("devtools") 
+# after that, say: 
+# devtools::install_github("benpeloquin7/rrrsa")
+# then you can load/attach `rrrsa::` to make its contents directly available: 
+library("rrrsa")
+
+
+# the package `lefftpack::` is a bundle of functions and data that i use 
+# frequently. the function `lazy_setup()` loads a handful of very useful 
+# packages (without startup messages), and also sets a `ggplot2::` theme 
+# with some convenient defaults. 
+# 
+# if you wanna use the package, you can install it just like `rrrsa::` above: 
+# devtools::install_github("lefft/lefftpack")
+lefftpack::lazy_setup()
+
+# alternatively, uncomment the next line to load some packages we'll use 
+# library("reshape2"); library("ggplot2"); library("magrittr"); library("dplyr")
+
+
+
+
+### CONTENTS OF `rrrsa::` PACKAGE ---------------------------------------------
+
+# SIX DATASETS
+
+# THREE HELPER FUNCTIONS
+
[…]

### 2017-10-20 paper01_peloquin_frank2016/model_walkthru/paper01_peloquin_frank2016.rmd
source code rewrite now done

@@
 ---
-title: "paper01 model walkthru"
-subtitle: "Peloquin & Frank (2016), 'Determining the alternatives for scalar implicature'"
+title: "Determining the alternatives for scalar implicature"
+subtitle: "Peloquin & Frank (2016)"
 author: "timothy leffel"
@@
 
-### Gameplan
+```{r}
+### ADAPTED CODE FROM PACKAGE VIGNETTE FOR DEMO OF fullRecursion()
+source("setup.r")
 
-- recap of paper 
-- what the model is (broad strokes)
-- what the model does 
-- ...
+# m/dat_pre ~~> matrix of semantics (rows=meaning (m rows), cols=words (n cols))
+# costs     ~~> ncol(m) vector of costs (default is 0 valued vector)
+# priors    ~~> nrow(m) vector of priors (default is uniform)
+# alpha     ~~> decision hyper-param
+dat_pre <- matrix(
+  nrow=5, ncol=3, dimnames=list(paste0("row", 1:5), c("none","some","all")), 
+  data=c(1.0, .00, .00, 
+         .00, .00, .00, 
+         .25, .25, .25, 
+         .25, .00, .00, 
+         .00, .00,  1)
+)
+dat_pre_df <- rsa_matrix_to_df(dat_pre, response_colname="semantics")
 
-### Recap of Paper
+costs <- c(none=0, some=0, all=0)
+priors <- rnorm(nrow(dat_pre), mean=.5, sd=.1)
+alpha <- 1
 
-### Model
+# rsa.fullRecursion(m=m, costs=rep(0, ncol(m)), priors=rep(1, nrow(m)), alpha=2)
+dat_post <- fullRecursion(m=dat_pre, 
+                          costs=rep(0,ncol(dat_pre)), 
+                          priors=rep(1,nrow(dat_pre)),
+                          alpha=alpha)
+
+knitr::kable(dat_pre, row.names=TRUE)
+knitr::kable(dat_post, row.names=TRUE)
+
+make_rsa_plot(dat_pre, dat_post)
+```
 
 ```{r}
+# source("setup.r")
+
+nams <- c("exp", "scale", "stars", "speaker.p", "words", "e11", "e6")
+#   1) quantityVarName :: "stars"
+#   2) semanticsVarName :: "speaker.p"
+#   3) itemVarName :: "words"
+#   4) groupVarName :: "scale"
+#   other) experiment number :: "exp"
+#   other) Pragmatic judgments study 1 (e6), study 2 (e11)
+sapply(rsa_data, dim)
+sapply(rsa_data, function(df){ identical(names(df), nams) })
+
+pf_data <- rsa_data[names(rsa_data)[names(rsa_data) != "frank"]]
+
+pf_data <- lapply(seq_along(pf_data), function(idx){
+  pf_data[[names(pf_data)[idx]]] %>% 
+    mutate(dataset = names(pf_data)[idx]) %>% 
+    mutate_if(is.factor, as.character) %>% 
+    rename(study1=e6, study2=e11)
+}) %>% 
+  (function(df_list) do.call(rbind, df_list))
+
+str(pf_data)
+# pf_data %>% filter(words=="none") %>% View() # ***
+
+# carve out the <none, some, all> scale (both expts 1 and 2) 
+# pf_data %>% filter(scale=="some_all") %>% View()
+pf3_NSA <- pf_data %>% 
+  filter(dataset=="pf3") %>% 
+  filter(scale=="some_all") %>% 
+  select(-exp, -scale, -dataset) %>% 
+  select(stars, words, study1, study2, speaker.p) %>% 
+  arrange(stars, words)
+
+# all response fields are 0 for none, across the whole scale...
+# pf3_NSA %>% filter(words=="none") %>% View() # ***
+
+pf3_NSA_study1 <- pf3_NSA %>% 
+  select(stars, words, value=study1) %>% 
+  tidyr::spread(words, value) %>% 
+  set_rownames(.$stars) %>%  # `stars` now encoded as rownames
+  select(none, some, all) %>% 
+  as.matrix()
+
+# this is the semantic representation (prob dist) of each scalar item 
+input_matrix <- pf3_NSA_study1
+
+# need to make sure length matches ncol
+costs <- c(none=0, some=0, all=0)
+
+# need to make sure length matches nrow 
+priors <- rnorm(nrow(input_matrix), mean=.5, sd=.1)
+
+# parameter that fixes "how much the cost matters" 
+alpha <- 1
+
+# calculate the posterior distributions for each scalar item 
+output_matrix <- fullRecursion(
+  input_matrix, costs=costs, priors=priors, alpha=alpha
+)
+
+knitr::kable(input_matrix, row.names=TRUE)
+knitr::kable(output_matrix, row.names=TRUE)
+
+make_rsa_plot(input_matrix, output_matrix)
+```
+
+```{r}
+# play around with all the params!!! 
+rsa.reason(m=input_matrix, 
+           costs=rep(0, ncol(input_matrix)), 
+           priors=rep(1, nrow(input_matrix)), 
+           depth=2, 
+           alpha=1, 
+           usePriorEveryRecurse=FALSE)
+
+reason(input_matrix, 
+       costs=rep(0, ncol(input_matrix)), 
+       priors=rep(1, nrow(input_matrix)), 
+       alpha=1, 
+       depth=2, 
+       recycle_priors=FALSE)
+
+rsa.fullRecursion(input_matrix, 
+                  costs=rep(0, ncol(input_matrix)), 
+                  priors=rep(1, nrow(input_matrix)),
+                  alpha=1)
+fullRecursion(input_matrix, 
+              costs=rep(0, ncol(input_matrix)), 
+              priors=rep(1, nrow(input_matrix)),
+              alpha=1)
+
+```
+
+```{r eval=FALSE, include=FALSE}
 lefftpack::lazy_setup()

### 2017-10-20 paper01_peloquin_frank2016/paper_and_materials/rrrsa_vignette.Rmd
source code rewrite now done

@@
 pragmaticsTidied <- res %>%
-  gather(word, pragmatics, -quantity)
+  tidyr::gather(word, pragmatics, -quantity)
 semanticsTidied <- as.data.frame(m) %>%
   mutate(quantity = rownames(.)) %>%
-  gather(word, semantics, c(none, some, all))
+  tidyr::gather(word, semantics, c(none, some, all))
 fullData <- merge(pragmaticsTidied, semanticsTidied) %>%

### 2017-10-20 paper01_peloquin_frank2016/model_walkthru/annotated-source-code/fullRecursion.r

@@ -1,0 +1,27 @@
+fullRecursion <- function(m, costs, priors, alpha=1){
+  
+  cost_mat <- matrix(
+    nrow=nrow(m), ncol=ncol(m), 
+    dimnames=list(paste0("row", 1:5), c("none","some","all")), 
+    data=rep(costs, times=1, each=nrow(m))
+  )
+  
+  likelihood <- t(mapply(utility, 
+                         items=split(m, f=rownames(m)), 
+                         costs=split(cost_mat, row(cost_mat)), 
+                         alpha=alpha))
+  # identical(cost_mat, costsAsMatrix)
+  
+  
+  unnorm_posterior <- apply(likelihood, MARGIN=2, function(col) priors * col)
+  
+  posterior <- apply(unnorm_posterior, MARGIN=2, normVec)
+  
+  dimnames(posterior) <- list(rownames(m), colnames(m))
+  
+  return(posterior)
+}
+
+
+
+

### 2017-10-20 paper01_peloquin_frank2016/model_walkthru/annotated-source-code/informativity.r

@@ -1,0 +1,17 @@
+
+informativity <- function(m_u, cost, alpha){
+  # `m_u` must be a probability, `alpha` a positive scalar 
+  inform <- ifelse(m_u == 0, 0, exp(-alpha * (-log(m_u) - cost)))
+  return(inform)
+}
+
+# m_us <- seq(from=0, to=1, by=.1)
+# alphas <- seq(from=1, to=10, by=2)
+# costs <- seq(from=0, to=5, by=1)
+# 
+# n <- 100
+# 
+# use combn() or outer() to put together
+# then plot it by varying each param 
+
+

### 2017-10-20 paper01_peloquin_frank2016/model_walkthru/annotated-source-code/normVec.r

@@ -1,0 +1,10 @@
+normVec <- function(v){
+  stopifnot(is.numeric(v))
+  
+  if (sum(v==0)==length(v)){
+    return(v)
+  } else {
+    # all elements of v must be positive or zero 
+    return(v/sum(v, na.rm=TRUE))
+  }
+}

### 2017-10-20 paper01_peloquin_frank2016/model_walkthru/annotated-source-code/reason.r

@@ -1,0 +1,16 @@
+# `rsa.reason()` is a wrapper function for `rsa.fullRecursion` which provides an additional `depth` parameter specifying the recursive depth during reasoning. If depth is not provided, default value is $1$.
+reason <- function(m, costs, priors, alpha, depth, recycle_priors){
+  depth_counter <- depth
+  while (depth_counter > 0){
+    if (!recycle_priors & depth_counter!=1){
+      m <- fullRecursion(m, costs=costs, priors=rep(1,nrow(m)), alpha=alpha)
+    } else {
+      m <- fullRecursion(m, costs=costs, priors=priors, alpha=alpha)
+    }
+    depth_counter <- depth_counter - 1
+  }
+  return(m)
+}
+
+
+

### 2017-10-20 paper01_peloquin_frank2016/model_walkthru/annotated-source-code/runDf.r

@@ -1,0 +1,84 @@
+
+### NEED TO FINISH THIS ONE UP + THEN EVERYTHING IS RE-WRITTEN 
+### NEED TO FINISH THIS ONE UP + THEN EVERYTHING IS RE-WRITTEN 
+### NEED TO FINISH THIS ONE UP + THEN EVERYTHING IS RE-WRITTEN 
+runDf <- function(data, quantityVarName, semanticsVarName, itemVarName, 
+                  costsVarName = NA, priorsVarName = NA, depth = 1, alpha = 1, 
+                  usePriorEveryRecurse = FALSE){
+  
+  # `c(quantityVarName, semanticsVarName, itemVarName)` must be in names(data)
+  # `semanticsVarName` needs to be numeric 
+  # need to specify costs column costsVarName in data
+  # same thing for priors column `priorsVarName`
+  
+  originalData <- data
+  
+  matrixLabels <- c(quantityVarName, semanticsVarName, itemVarName)
+  
+  matrixData <- data %>% 
+    dplyr::select_(quantityVarName, semanticsVarName, itemVarName) %>% 
+    tidyr::spread_(itemVarName, semanticsVarName) %>% 
+    as.data.frame()
+  
+  rownames(matrixData) <- matrixData[[quantityVarName]]
+  
+  matrixData <- matrixData %>% dplyr::select(-1) %>% data.matrix()
+  
+  if (is.na(costsVarName)){
+    costs <- rep(0, length(unique(data[[itemVarName]])))
+    names(costs) <- unique(data[[itemVarName]])
+  } else {
+    
+    costsData <- data %>% 
+      dplyr::select_(itemVarName, costsVarName) %>% unique()
+    
+    costs <- costsData[[costsVarName]]
+    
+    names(costs) <- costsData[[itemVarName]]
+  }
+  
+  if (is.na(priorsVarName)){
+    priors <- rep(1, length(unique(data[[quantityVarName]])))
+    
+    names(priors) <- unique(data[[quantityVarName]])
+  } else {
+    
+    priorsData <- data %>% 
+      dplyr::select_(quantityVarName, priorsVarName) %>% unique()
+    
+    priors <- priorsData[[priorsVarName]]
+    
+    names(priors) <- priorsData[[quantityVarName]]
+    
+    quantityVec <- priorsData[[quantityVarName]]
+  }
+  
+  posteriors <- rsa.reason(
[…]

### 2017-10-20 paper01_peloquin_frank2016/model_walkthru/annotated-source-code/scratch.r

@@ -1,0 +1,154 @@
+
+
+dat_post_df <- rsa_matrix_to_df(dat_post, response_colname="pragmatics")
+
+dat_prepost <- full_join(dat_pre_df, dat_post_df, by=c("quantity", "word")) %>% 
+  melt(id.vars=c("quantity", "word")) %>% 
+  rename(type = variable) %>% 
+  mutate(type = as.character(type))
+
+ggplot(dat_prepost, aes(x=quantity, y=value, col=word)) + 
+  geom_line() + facet_wrap(~type)
+
+
+
+
+dat_pre_df <- as.data.frame(dat_pre) %>% 
+  mutate(quantity = as.numeric(gsub("row", "", rownames(dat_pre)))) %>% 
+  melt(id.vars="quantity", variable.name="word") %>% 
+  mutate(word = as.character(word)) %>% 
+  rename(semantics = "value") %>% 
+  arrange(quantity, word)
+
+
+
+
+
+# d <- rsa_data$pf2
+# d %>% select(scale, words, stars, expt1=e6, expt2=e11, speaker.p) %>% 
+#   mutate(scale = gsub("ellent", "", scale)) %>% 
+#   mutate(words = gsub("ellent", "", words)) %>% head
+
+
+
+
+
+
+
+
+
+m
+row(m)
+(xx <- matrix(c(1,2,3,4,5,6,4,4), ncol=2))
+# both same shape, row has idx along cols; col has idx along rows 
+row(xx); col(xx) 
+# 
+split(xx, f=xx[,1])
+split(xx, f=xx[,2])
+
+split(m, f=rownames(m))
+split(costsAsMatrix, row(costsAsMatrix))
+
+
+mapply(rep, 1:4, 4:1)
+mapply(rep, times = 1:4, x = 4:1)
+
+mapply(rep, 1:4, times=4:1, each=4:1)
+
+
+mapply(rep, times = 1:4, MoreArgs = list(x = 42))
+mapply(rep, times = 1:4, MoreArgs = list(x = 42, each=2))
+
+mapply(function(x, y) seq_len(x) + y,
+       c(a =  1, b = 2, c = 3),  # names from first
+       10)#c(A = 10, B = 0, C = -10))
+
+
+
+word <- function(C, k) paste(rep.int(C, k), collapse = "")
+utils::str(mapply(word, LETTERS[1:6], 6:1, SIMPLIFY = FALSE))
+
+
+
[…]

### 2017-10-20 paper01_peloquin_frank2016/model_walkthru/annotated-source-code/tuneDepthAlpha.r

@@ -1,0 +1,71 @@
+
+
+
+# function that ...
+rsa.tuneDepthAlpha <- function(
+  data, quantityVarName, semanticsVarName, itemVarName, 
+  groupName = NA, compareDataName, costsVarName = NA, priorsVarName = NA, 
+  depths = 1, alphas = 1, compareIndices = NA, usePriorEveryRecurse = TRUE
+){
+  # initialize empty df object 
+  cors <- data.frame(cor = NA, depth = NA, alpha = NA)
+
+  # if `groupName` is NA, then nothing happens and empty df is returned 
+  if (!is.na(groupName)) {
+
+    # for each `a` in `alphas` 
+    for (a in alphas) {
+
+      # and each `d` in `depths` 
+      for (d in depths) {
+
+        # execute an RSA iteration on `data` with `depth=d` and `alpha=a` 
+        # (can also do this by-groups, if `groupName` is specified)
+        currRun <- plyr::ddply(data, .fun = rsa.runDf, 
+                               .variables = c(groupName), 
+                               quantityVarName = quantityVarName, 
+                               semanticsVarName = semanticsVarName, 
+                               itemVarName = itemVarName, 
+                               costsVarName = costsVarName, 
+                               priorsVarName = priorsVarName, 
+                               depth = d, alpha = a, 
+                               usePriorEveryRecurse = usePriorEveryRecurse)
+
+        # if `compareIndices` is NA, then compute correlation between 
+        # entire `compareDataName` and model predictions for the current run 
+        # (also recording `a` and `d` for inspection) 
+        if (length(compareIndices) == 1 & is.na(compareIndices[1])) {
+          res <- c(
+            cor = cor(currRun[, compareDataName], currRun[, "preds"], 
+                      use="pairwise.complete.obs"), 
+            depth = d, 
+            alpha = a
+          )
+
+          # add the correlation for this run to the correlations df
+          cors <- rbind(cors, res)
+
+        } else {
+          # if `compareIndices` are specified, then cut the model output df 
+          # to just those indices, then compute + record the correlation coef
+          compareData <- currRun[compareIndices, ]
+          res <- c(
+            cor = cor(compareData[, compareDataName], compareData[, "preds"], 
+                      use = "pairwise.complete.obs"), 
+            depth = d, 
+            alpha = a
+          ) 
+          # and add the coef to the correlations df 
+          cors <- rbind(cors, res)
+        }
[…]

### 2017-10-20 paper01_peloquin_frank2016/model_walkthru/annotated-source-code/utility.r

@@ -1,0 +1,38 @@
+utility <- function(items, costs, alpha){
+  # costs must have same length as items, alpha must be a positive scalar 
+  
+  # apply `informativity()` to each item with specified costs and alpha. 
+  # then collect them all and return with the same shape/order as `items`
+  inform <- mapply(informativity, m_u=items, cost=costs, alpha=alpha)
+  
+  # normalization here is just dividing each number by the sum of them all 
+  inform_norm <- normVec(inform)
+  
+  return(inform_norm)
+}
+
+
+
+# ### ORIGINAL 
+# split(m, row(m))
+# split(cost_mat, row(cost_mat))
+# split(costsAsMatrix, row(costsAsMatrix))
+# alpha=alpha
+# 
+# (function (items, costs = rep(0, length(items)), alpha = 1) 
+# {
+#   rsa.normVec(mapply(rsa.informativity, items, costs, alpha = alpha))
+#   
+#   
+# })(split(m, row(m))[[1]], split(costsAsMatrix, row(costsAsMatrix))[[1]], alpha=alpha)
+# 
+# utility(split(m, row(m))[[1]], split(costsAsMatrix, row(costsAsMatrix))[[1]], alpha=alpha)
+# 
+# utility(split(m, row(m))[[1]], split(costsAsMatrix, row(costsAsMatrix))[[1]], alpha=alpha)
+# 
+# utility(items=split(m, row(m))[[1]], costs=split(costsAsMatrix, row(costsAsMatrix))[[1]], alpha=alpha)
+# 
+# rrrsa::rsa.utility(items=split(m, row(m))[[1]], costs=split(costsAsMatrix, row(costsAsMatrix))[[1]], alpha=alpha)
+
+
+# (split(m, row(m))[[1]], split(m, row(m))[[1]], alpha=alpha)

### 2017-10-20 paper01_peloquin_frank2016/model_walkthru/notes.r

@@ -38,7 +38,7 @@ library("rrrsa")
 # devtools::install_github("lefft/lefftpack")
 lefftpack::lazy_setup()
 
-# alternatively, uncomment the next line to load some packages we'll use 
+# alternatively, uncomment the next line to load the packages we'll use 
 # library("reshape2"); library("ggplot2"); library("magrittr"); library("dplyr")
 
 
@@ -46,12 +46,42 @@ lefftpack::lazy_setup()
 
 ### CONTENTS OF `rrrsa::` PACKAGE ---------------------------------------------
 
-# SIX DATASETS
+# FIVE DATASETS
+rsa_data <- list(
+  # from frank et al submitted
+  frank = rrrsa::d_pragmods,
+  # from pelfrank16 
+  pf2   = peloquinFrank_2Alts,
+  pf3   = peloquinFrank_3Alts,
+  pf4   = peloquinFrank_4Alts,
+  pf5   = peloquinFrank_5Alts
+)
 
 # THREE HELPER FUNCTIONS
+rsa_helper <- list(
+  # converts vec2 to whatever type vec1 is
+  convertVecType = rrrsa::rsa.convertVecType, 
+  # this is just `function(x) x/sum(x)` for positive `x`
+  normVec        = rrrsa::rsa.normVec, 
+  # just changes desired column names 
+  renameCol      = rrrsa::rsa.renameCol
+)
 
 # SIX ANALYSIS FUNCTIONS
-
+rsa_model <- list(
+  # computes informativity given params `m_u`, `alpha`, `cost`
+  informativity  = rrrsa::rsa.informativity,
+  # computes utility given params `items`, `costs`, `alpha`
+  utility        = rrrsa::rsa.utility,
+  # wrapper that runs `depth`-many iterations of `rsa.fullRecursion()` 
+  reason         = rrrsa::rsa.reason,
+  # wrapper around `rsa.reason` that accepts input as a data frame 
+  runDf          = rrrsa::rsa.runDf,
+  # explores correlation between data and model preds for various of `alpha` 
+  tuneDepthAlpha = rrrsa::rsa.tuneDepthAlpha,
+  # main function implementing RSA model 
+  fullRecursion  = rrrsa::rsa.fullRecursion
+)
 
 
 
@@ -149,7 +179,7 @@ rrrsa::rsa.informativity(m_u, alpha, cost)
[…]

### 2017-10-20 paper01_peloquin_frank2016/model_walkthru/setup.r

@@ -1,0 +1,94 @@
+lefftpack::lazy_setup()
+library("rrrsa")
+
+source("annotated-source-code/normVec.r")
+source("annotated-source-code/utility.r")
+source("annotated-source-code/informativity.r")
+source("annotated-source-code/fullRecursion.r")
+source("annotated-source-code/reason.r")  
+# source("annotated-source-code/runDf.r") # DOES NOT EXISST HYETS!
+source("annotated-source-code/tuneDepthAlpha.r")
+
+
+### CONTENTS OF `rrrsa::` PACKAGE ---------------------------------------------
+
+# FIVE DATASETS
+rsa_data <- list(
+  # from frank et al submitted
+  frank = d_pragmods,
+  # from pelfrank16 
+  pf2   = peloquinFrank_2Alts,
+  pf3   = peloquinFrank_3Alts,
+  pf4   = peloquinFrank_4Alts,
+  pf5   = peloquinFrank_5Alts
+)
+
+# THREE HELPER FUNCTIONS
+rsa_helper <- list(
+  # converts vec2 to whatever type vec1 is
+  convertVecType = rrrsa::rsa.convertVecType, 
+  # this is just `function(x) x/sum(x)` for positive `x`
+  normVec        = rrrsa::rsa.normVec, 
+  # just changes desired column names 
+  renameCol      = rrrsa::rsa.renameCol
+)
+
+# SIX ANALYSIS FUNCTIONS
+rsa_model <- list(
+  # computes informativity given params `m_u`, `alpha`, `cost`
+  informativity  = rrrsa::rsa.informativity,
+  # computes utility given params `items`, `costs`, `alpha`
+  utility        = rrrsa::rsa.utility,
+  # wrapper that runs `depth`-many iterations of `rsa.fullRecursion()` 
+  reason         = rrrsa::rsa.reason,
+  # wrapper around `rsa.reason` that accepts input as a data frame 
+  runDf          = rrrsa::rsa.runDf,
+  # explores correlation between data and model preds for various of `alpha` 
+  tuneDepthAlpha = rrrsa::rsa.tuneDepthAlpha,
+  # main function implementing RSA model 
+  fullRecursion  = rrrsa::rsa.fullRecursion
+)
+
+
+column_key <- setNames(
+  object=c("stars","speaker.p","words","scale","expt1","expt2"), 
+  nm=c("quantityVarName", "semanticsVarName", "itemVarName", 
[…]

### 2017-10-20 paper01_peloquin_frank2016/model_walkthru/paper01_peloquin_frank2016.rmd
push pre-final version in case anyone is working on this before the meeting

@@
 ---
-title: "Determining the alternatives for scalar implicature"
-subtitle: "Peloquin & Frank (2016)"
-author: "timothy leffel"
-date: "oct20/2017"
+title: "Model Walkthru for Peloquin & Frank (2016)"
+subtitle: "\"Determining the alternatives for scalar implicature\""
+author: "timothy leffel, oct20/2017"
+output: 
+  html_document:
+    css: style.css
+    self_contained: no
 ---
+
+```{r setup}
+source("setup.r")
+```
+
+## Quick recap of paper
+
+## Contents of model implementation package
+
+Below is a listing of the contents of the `rrrsa::` package.
+
+#### Five datasets 
+```{r}
+# FIVE DATASETS
+rsa_data <- list(
+  # from frank et al submitted:
+  frank = rrrsa::d_pragmods,
+  # from pelfrank16: 
+  pf2   = rrrsa::peloquinFrank_2Alts,
+  pf3   = rrrsa::peloquinFrank_3Alts,
+  pf4   = rrrsa::peloquinFrank_4Alts,
+  pf5   = rrrsa::peloquinFrank_5Alts
+)
+```
+
+#### Three helper functions 
+```{r}
+# THREE HELPER FUNCTIONS
+rsa_helper <- list(
+  # converts vec2 to whatever type vec1 is
+  convertVecType = rrrsa::rsa.convertVecType, 
+  # this is just `function(x) x/sum(x)` for positive `x`
+  normVec        = rrrsa::rsa.normVec, 
+  # just changes desired column names 
+  renameCol      = rrrsa::rsa.renameCol
+)
+```
+
+#### Six analysis functions 
+```{r}
+# SIX ANALYSIS FUNCTIONS
+rsa_model <- list(
+  # computes informativity given params `m_u`, `alpha`, `cost`
+  informativity  = rrrsa::rsa.informativity,
+  # computes utility given params `items`, `costs`, `alpha`
+  utility        = rrrsa::rsa.utility,
+  # wrapper that runs `depth`-many iterations of `rsa.fullRecursion()` 
+  reason         = rrrsa::rsa.reason,
+  # wrapper around `rsa.reason` that accepts input as a data frame 
+  runDf          = rrrsa::rsa.runDf,
+  # explores correlation between data and model preds for varying `alpha` 
+  tuneDepthAlpha = rrrsa::rsa.tuneDepthAlpha,
+  # main function implementing RSA model 
+  fullRecursion  = rrrsa::rsa.fullRecursion
+)
+```
+
+## An example in action
 
@@
 ### ADAPTED CODE FROM PACKAGE VIGNETTE FOR DEMO OF fullRecursion()
-source("setup.r")
 
@@
 
+## Reproducing the paper results
+
+#### get acquainted with the datasets, combine them
+
 ```{r}
-# source("setup.r")
+source("setup.r")
 
@@
 # pf_data %>% filter(words=="none") %>% View() # *** 
+```
 
+#### Specific example: Expt1, three alts, <*none, some, all*>
+
+```{r}
 # carve out the <none, some, all> scale (both expts 1 and 2) 
@@
 
+#### Specific example: Expt2, three alts, <*none, some, all*>
+
+#### Specific example: ...
+
+## Play around with parameters
+
 ```{r}
@@
 
+## Concepts used in RSA models
+
+#### primitive principles
+
+- speech-acts are rational behaviors -- actions in **signaling games** 
+- utterances are selected by their **expected utility** 
+- the utility of an utterance is a function of its **informativity** and its **cost**
+
+#### model of a conversational exchange
+
+- a **speaker** $S_n$ wants to communicate some proposition 
+- $S_n$ selects utterance $u$ from a set of alternative utterances 
+- an actual ("pragmatic") **listener** ???
+- FINISH
+
+## The mathematics underlying RSA models
+
+The following equations are from Goodman & Frank's (2012) model specification.
+
+#### equation 1 (GF2012)
+
+$$P(r_s|w, C) = \frac{P(w|r_s, C)\times P(r_s)}{\sum\limits_{r'\in C} P(w|r', C) \times P(r')}$$
+
+where
+
+- $r_s$ is an intended referent (element of $C$), 
+- $C$ is a context (set of objects $r, r', ...\in C$), and 
+- $w$ is the word uttered by the speaker. 
+- The **prior** is $P(r_s)$, 
+- the **likelihood** is $P(w|r_s, C)$, and 
+- the **normalizing constant** is $\sum\limits_{r'\in C} P(w|r', C) \times P(r')$
+
+#### equation 2 (GF2012)
+
+$$P(w|r_s, C) = \frac{|w|^{-1}}{\sum\limits_{w'\in W} |w'|^{-1}}$$
+
+where 
+- $|w|$ is the number of objects in $C$ to which word $w$ could(?) apply, 
+- and $W$ is the set of words that apply to the intended referent $r_s$.
+
+##### equation S1 (GF12 supplement)
+
+$$P(w|r_s, C) \propto e^{\alpha\times U(w; r_s, C)}$$
+
+where
+
+- the **context** $C = \{o_1, ..., o_n\}$ is a set of objects 
+- the **vocabulary** is a set of words $V = \{w_1, ..., w_m\}$ 
+- each word $w\in V$ has a meaning $[\![ w ]\!] : C \rightarrow \{0, 1\}$ 
+- words are chosen in proportion to their expected utility
+
+#### equation S2 -- utility (GF12 supplement)
+
+$$U(w; r_s, C) = I(w; r_s, C) - D(w)$$
+
+where
+
+- $I(w; r_s, C)$ is the informativeness of utterance $w$ w.r.t. intended referent $r_s \in C$, and 
+- $D(w)$ is $w$'s cost (in $C$?). (for single words, $D : V \rightarrow \mathbb{R}^+$ is constant)
+
+##### surprisal (GF12 supplement)
+
+for a sample $x$ from a known distribution $p(x)$, the surprisal of $x$ is
+$$I_p(x) = -log(p(x))$$
+
+##### utility and surprisal (GF12 supplement)
+
+utility decreases with surprisal:
+
+$$I(w; r_s, C) = -I_{\widetilde{w}_C}(r_s)$$
+
+where $\widetilde{w}_C$ is the distribution over objects that would come froma  aliteral interpretation of $w$ in context $C$.
+
+##### equation S3 -- literal listener model derivation (GF12, supplementary materials)
+
+"if listeners interpret the utterance $w$ literally, assigning zero probability to objects for which the word is false, they assign equal probability to each object conssitent with $w$.
+This distribution over objects can be written:"
+
+$$\widetilde{w}_C(o) = 
+    \begin{cases}
+    \frac{1}{|w|} & \text{if } w(o) = true\\\\
+    0             & \text{otherwise }
+    \end{cases}
+$$
+
+##### equation S4 -- listener likelihood (GF12, supplement)
+
+equation S4 follows from S1-S3, which is equivalent to equation 2 (aka "the size principle")
+
+$$P(w|r_s, C) = \frac
+                    {e^{-(-\log(|w|^{-1}))}}
+                    {\sum\limits_{w'\in V \text{s.t. } w'(r_s) = true} e^{-(-\log(|w'|^{-1}))}}$$
+
+"Thus in our experiments, the speaker's abstract goal of being informative reduces to a simple formulation: whoose a word that applies to the referent and picks out a relatively smaller section of the context.
+Listeners may then use this model of a speaker as their likelihood function, to be combined with prior information about contextual salience as in Equation 1 in the main text."
+
+<!-- A `pragmatic listener` $P_{L_n}(m|u)$, reasons about intended meaning $m$ of an utterance $u$ by a `rational speaker` $P_{s_n}(u|m)$ who chooses an utterance according to the expected utility of an utterance $U(m;u)$.
+$\alpha$ is a decision noise parameter.
+-->
+
+<!-- $$P_{L_n}(m|u) \propto P_{S_n}(u|m)P(m)$$ -->
+<!-- $$P_{S_n} \propto e^{U(m;u)}$$ -->
+<!-- $$U(m;u) = -\alpha(-\log(P_{L_{n-1}}(m|u)) - C(u))$$ -->
+
+<!-- See Frank & Goodman (2012) and Goodman & Stuhmuller (2013) for the original descriptions of the RSA framework.
+Frank, et al.
+(Under Review) also provides a comprehensive presentation and evaluation of RSA.
+-->
+
 ```{r eval=FALSE, include=FALSE}
-lefftpack::lazy_setup()
-# devtools::install_github("benpeloquin7/rrrsa")
-library("rrrsa")
-
-### CONTENTS -------
-
-# SIX DATASETS
-
-# THREE HELPER FUNCTIONS
-
-# SIX ANALYSIS FUNCTIONS
-
-### DATASETS ------------------------------------------------------------------
-
-# note: on any dataset `df`, start by calling `?df` to see info
-
-# example docs for `rrrsa::peloquinFrank_2Alts`
-# 
-#   This data set contains empirical literal listener semantics and pragmatic
-#   judgments for entailment items (two alternatives) from Peloquin & Frank 2016
-# 
-#   1) quantityVarName :: "stars"
-#   2) semanticsVarName :: "speaker.p"
-#   3) itemVarName :: "words"
-#   4) groupVarName :: "scale"
-#   other) experiment number :: "exp"
-#   other) Pragmatic judgments study 1 (e6), study 2 (e11)
-# 
-#   see also: https://github.com/langcog/scalar_implicature
-
-# 136x16 df, data from frank et al. (under review)
-rrrsa::d_pragmods
-
-# 50x7 df, data from pel+frk16, 
-rrrsa::peloquinFrank_2Alts
-
-# 75x7 df
-rrrsa::peloquinFrank_3Alts
-
-# 100x7 df
-rrrsa::peloquinFrank_4Alts
-
-# 125x7 df 
-rrrsa::peloquinFrank_5Alts
-
-### HELPER FUNCTIONS ----------------------------------------------------------
-
-# converts vec2 to whatever type vec1 is
-rrrsa::rsa.convertVecType(vec1, vec2)
-
-# this is just `function(x) x/sum(x)` for positive `x`
-rrrsa::rsa.normVec(v)
-
-# just changes desired column names 
-rrrsa::rsa.renameCol(df, currNames, replacements)
-
-### ANALYSIS FUNCTIONS --------------------------------------------------------
-
-### `rrrsa::rsa.informativity(m_u, alpha, cost)` -------
-# ifelse(m_u == 0, 0, exp(-alpha * (-log(m_u) - cost)))
-rrrsa::rsa.informativity(m_u, alpha, cost)
-
-### `rrrsa::rsa.utility(items, costs, alpha)` ------
-# rsa.normVec(mapply(rsa.informativity, items, costs, alpha = alpha))
-rrrsa::rsa.utility(items, costs, alpha)
-
-### `rrrsa::rsa.reason(m, costs, priors, depth, alpha)` ----
-rrrsa::rsa.reason(m, costs, priors, depth, alpha, usePriorEveryRecurse)
-
-### `rrrsa::rsa.runDf()` ----------------
-rrrsa::rsa.runDf(data, quantityVarName, semanticsVarName, 
-                 itemVarName, costsVarName, priorsVarName, depth, alpha,
-                 usePriorEveryRecurse)
-
-### `rrrsa::rsa.tuneDepthAlpha()` --------
-rrrsa::rsa.tuneDepthAlpha()
-
 ### rrrsa::rsa.fullRecursion(m, costs, priors, alpha) ------

### 2017-10-20 paper01_peloquin_frank2016/model_walkthru/paper01_peloquin_frank2016.rmd
ready for walkthru

@@
 author: "timothy leffel, oct20/2017"
-output: 
-  html_document:
-    css: style.css
-    self_contained: no
+output: html_document
 ---
@@
 ```{r setup}
+# this loads all the rsa functions, rewritten for ease of reading 
 source("setup.r")
@@
 
-## Quick recap of paper
+## 1. Recap of paper
 
-## Contents of model implementation package
+#### Experiment 2: Literal Listener Task
+
+> Participants were presented with a target scalar item and a star rating (1–5 stars) and asked to judge the compat- ibility of the scalar item and star rating.
+Compatibility was assessed through a binary “yes/no” response to a question of the form, “Do you think that the person thought the food was ____ ?” where a target scalar was presented in the blank.
+Each participant saw all scalar item and star rating combinations for their particular condition, in a random order.
+<br><br> The two alternatives condition only included entailment pairs taken from van Tiel et al.
+(2014) for a total of 50 trials for each participant.
+The four alternatives condition included entailment pairs plus the top two alternatives generated for each scalar family by participants in Experiment 1 (100 trials per participant).
+The five alternatives condition included the four previous items plus one more neutral item chosen from alternatives generated in Experiment 1 (125 trials per participant).
+[SEE FIGURE 3, LEFT]
+
+```{r out.width="75%"}
+knitr::include_graphics("input/pf16_fig3_left.png")
+```
+
+> Literal listener judgments from Experiment 2.
+Proportion of participants indicating compatibility (answering “yes”) is shown on the vertical axis, with the horizontal axis showing number of stars on which the utterance was judged.
+Rows are grouped by scale and items within rows are ordered by valence.
+Colors indicate the specific condition with conditions including different numbers of items.
+
+```{r}
+knitr::include_graphics("input/pf16_fig4.png")
+```
+
+#### Experiment 3: Pragmatic Listener Task
+
+##### Setup
+
+> Participants were presented with a one-sentence prompt such as “Someone said they thought the food was ____ .” with a target scalar item in the blank.
+Partic- ipants were then asked to generate a star rating representing the rating they thought the reviewer likely gave.
+Each participant was presented with all scalar items in a random order.
+Participants in the two alternatives condition gave a total of 10 pragmatic judgments.
+Participants in the four alternatives condition gave a total of 20 pragmatic judgments.
+[SEE FIGURE 3, RIGHT]
+
+```{r out.width="75%"}
+knitr::include_graphics("input/pf16_fig3_right.png")
+```
+
+##### Results
+
+> Pragmatic listener judgments from Experiment 3.
+Vertical axis shows proportion of participants generating a star rating.
+Horizontal axis shows number of stars on which the utterance was judged.
+Line type denotes condition and colors indicate the particular scalar items.
+Each panel shows one entailment scalar pair.
+
+```{r}
+knitr::include_graphics("input/pf16_fig5.png")
+```
+
+#### Model Simulations
+
+##### Setup
+
+> Using literal listener data from Experiment 2, we conducted a set of simulations with the RSA model.
+Each simulation kept the model constant, fitting the choice parameter α as a free parameter, but used a set of alternatives to specify the scale over which predictions were computed.
+We considered four different alternative sets, with empirical measurements corresponding to those shown in Figure 1: 1) the two alternatives in the classic entailment scales, 2) those two alternatives with the addition of a generic negative alternative, 3) the expanded set of four alternatives, and 4) the expanded set of five alter- natives.
+Literal semantics for the generic negative alternative served as a baseline “none” semantics in which the scalar item was only compatible with 1 star.
+[SEE FIGURE 1]
+
+```{r out.width="75%"}
+knitr::include_graphics("input/pf16_fig1.png")
+```
+
+##### Results
+
+> Model fit with human judgments was significantly improved by the inclusion of alternatives beyond the entailment items (Table 1).
+The two alternatives model contained only entailment items, which, under classic accounts, should be sufficient to generate implicature, but fit to data was quite poor with these items.
+The addition of a generic negative element produced some gains, but much higher performance was found when we included four and five alternatives, with the alternatives derived empirically for the specific scale we used.
+[SEE FIGURE 6]
+
+```{r}
+knitr::include_graphics("input/pf16_fig6.png")
+```
+
+```{r}
+knitr::include_graphics("input/pf16_table1.png")
+```
+
+## 2. Overview of model implementation code
 
@@
 
-## An example in action
+## 3. Reproduce the paper results
 
+### 3.1 A toy example (adapted from package vignette)
+
+#### Get the literal semantics for *none*, *some*, *all* 
 ```{r}
-### ADAPTED CODE FROM PACKAGE VIGNETTE FOR DEMO OF fullRecursion()
-
-# m/dat_pre ~~> matrix of semantics (rows=meaning (m rows), cols=words (n cols))
-# costs     ~~> ncol(m) vector of costs (default is 0 valued vector)
-# priors    ~~> nrow(m) vector of priors (default is uniform)
-# alpha     ~~> decision hyper-param
-dat_pre <- matrix(
+semantics_matrix <- matrix(
   nrow=5, ncol=3, dimnames=list(paste0("row", 1:5), c("none","some","all")), 
@@
 )
-dat_pre_df <- rsa_matrix_to_df(dat_pre, response_colname="semantics")
+# semantics_df <- rsa_matrix_to_df(semantics_matrix, 
+#                                  response_colname="semantics")
+```
+
+#### Fix model parameters 
+```{r}
+### NEED TO MAKE SURE THESE MATCH THE SHAPE OF THE SEMANTICS MATRIX 
+# m/dat_pre ~~> matrix of semantics (rows=meaning (m rows), cols=words (n cols))
+# costs     ~~> ncol(m) vector of costs (default is 0 valued vector)
+# priors    ~~> nrow(m) vector of priors (default is uniform)
+# alpha     ~~> decision hyper-param
 
 costs <- c(none=0, some=0, all=0)
-priors <- rnorm(nrow(dat_pre), mean=.5, sd=.1)
+priors <- rnorm(nrow(semantics_matrix), mean=.5, sd=.1)
 alpha <- 1
 
-# rsa.fullRecursion(m=m, costs=rep(0, ncol(m)), priors=rep(1, nrow(m)), alpha=2)
-dat_post <- fullRecursion(m=dat_pre, 
-                          costs=rep(0,ncol(dat_pre)), 
-                          priors=rep(1,nrow(dat_pre)),
-                          alpha=alpha)
-
-knitr::kable(dat_pre, row.names=TRUE)
-knitr::kable(dat_post, row.names=TRUE)
-
-make_rsa_plot(dat_pre, dat_post)
+print(costs)
+print(priors)
 ```
 
-## Reproducing the paper results
+#### Calculate posterior 
+```{r}
+# same as: 
+# rsa.fullRecursion(m=m, costs=rep(0, ncol(m)), priors=rep(1, nrow(m)), alpha=2)
+pragmatics_matrix <- fullRecursion(
+  m=semantics_matrix, 
+  costs=costs, priors=priors, alpha=alpha
+)
+# pragmatics_df <- rsa_matrix_to_df(semantics_matrix,
+#                                   response_colname="pragmatics")
+
+knitr::kable(semantics_matrix, row.names=TRUE)
+knitr::kable(pragmatics_matrix, row.names=TRUE)
+```
+
+#### Visualize "semantics" versus "pragmatics"
+
+```{r}
+make_rsa_plot(semantics_matrix, pragmatics_matrix)
+```
+
+### 3.2 Inspect actual experimental data
 
@@
 ```{r}
-source("setup.r")
+# each dataset from the paper has these columns: 
+nams <- c("exp", "scale", "stars", "speaker.p", "words", "e11", "e6")
 
-nams <- c("exp", "scale", "stars", "speaker.p", "words", "e11", "e6")
+# column descriptions from documentation: 
+# 
 #   1) quantityVarName :: "stars"
@@
 #   other) Pragmatic judgments study 1 (e6), study 2 (e11)
+
+# dimensions and colnames of each dataset 
 sapply(rsa_data, dim)
@@
 
+# just want to consider the data from the peloquin & frank paper 
 pf_data <- rsa_data[names(rsa_data)[names(rsa_data) != "frank"]]
 
+# combine all of the PF16 datasets into a single df 
 pf_data <- lapply(seq_along(pf_data), function(idx){
@@
 
+# and look at its structure
 str(pf_data)
-# pf_data %>% filter(words=="none") %>% View() # *** 
+
+# each row is a unique combo of scale, stars, words, dataset 
+pf_data %>% group_by(scale, stars, words, dataset) %>% count() %$% table(n)
 ```
 
-#### Specific example: Expt1, three alts, <*none, some, all*>
+#### Reproduce Figure 4 from PF16
+
+NOTE: these are by-item averages for Expt2 (the number of trials for 2, 4, and 5 alternatives matches description on page 321)
 
 ```{r}
-# carve out the <none, some, all> scale (both expts 1 and 2) 
+d <- pf_data %>% 
+  filter(dataset %in% c("pf2", "pf4", "pf5"))
+
+knitr::kable(d %>% group_by(dataset) %>% summarize(rows = n()))
+```
+
+Here are the words used for each scale in Expt2: 
+```{r}
+d %>% group_by(dataset, scale) %>% summarize(
+  rows = n(), 
+  words = paste(sort(unique(words)), collapse=" | ")
+) %>% knitr::kable()
+```
+
+Unclear which column is plotted in Figure 4, so try all three and inspect results of each.
+
+```{r}
+d %>% 
+  ggplot(aes(x=stars, y=speaker.p, color=scale, linetype=dataset)) + 
+  geom_line() + 
+  facet_wrap(~words) + 
+  labs(title="fig4(?) -- plotting column d$speaker.p")
+```
+
+```{r}
+d %>% 
+  ggplot(aes(x=stars, y=study1, color=scale, linetype=dataset)) + 
+  geom_line() + 
+  facet_wrap(~words) + 
+  labs(title="fig4(?) -- plotting column d$study1")
+```
+
+```{r}
+d %>% 
+  ggplot(aes(x=stars, y=study2, color=scale, linetype=dataset)) + 
+  geom_line() + 
+  facet_wrap(~words) + 
+  labs(title="fig4(?) -- plotting column d$study2")
+```
+
+#### Reproduce Figure 5 from PF16
+
+```{r}
+lo_words <- c("good","liked","memorable","palatable","some")
+hi_words <- c("excellent","loved","unforgettable","delicious","all")
+
+d <- pf_data %>% 
+  filter(dataset %in% c("pf2", "pf4")) %>% 
+  filter(words %in% c(lo_words, hi_words))
+
+table(d$dataset) # color is 4alts, 2alts
+table(d$scale)   # facet by scale 
+table(d$words)
+
+d$scale_pos <- case_when(
+  d$words %in% lo_words ~ "lo", 
+  d$words %in% hi_words ~ "hi"
+)
+
+# two rows per combo, from `$exp` column (e8 and e10) 
+d_summary <- d %>% group_by(scale, scale_pos, stars) %>% summarize(
+  rows = n(), 
+  mean_speaker_p = mean(speaker.p),
+  mean_study1 = mean(study1), 
+  mean_study2 = mean(study2)
+) %>% ungroup()
+
+d_summary %>% 
+  # also include  "mean_speaker_p" to show model predictions(????)
+  melt(measure.vars=c("mean_study1", "mean_study2")) %>%  
+  ggplot(aes(x=stars, y=value, color=variable)) + 
+  # geom_point() + 
+  geom_line(aes(linetype=scale_pos)) + 
+  facet_wrap(~scale, nrow=1) + 
+  labs(title="reproduction of figure 5 from PF16")
+
+d_summary %>% ggplot(aes(x=stars, y=mean_study1, color=scale_pos)) + 
+  geom_point() + 
+  geom_line() + 
+  facet_wrap(~scale)
+```
+
+### 3.3 Specific example: Expt2, three alts, <*none, some, all*>
+
+```{r}
+# carve out the <none, some, all> scale from three-alts(??)
 # pf_data %>% filter(scale=="some_all") %>% View()
@@
 alpha <- 1
+```
 
+#### Calculate posterior
+
+```{r}
 # calculate the posterior distributions for each scalar item 
@@
 
-#### Specific example: Expt2, three alts, <*none, some, all*>
-
-#### Specific example: ...
-
-## Play around with parameters
+### 3.3 Specific example: [WHATEVER SEEMS MOST INTERESING!]
 
 ```{r}
-# play around with all the params!!! 
-rsa.reason(m=input_matrix, 
-           costs=rep(0, ncol(input_matrix)), 
-           priors=rep(1, nrow(input_matrix)), 
-           depth=2, 
-           alpha=1, 
-           usePriorEveryRecurse=FALSE)
+### STEPS FOR DOING ONE FULL ITERATION OF RSA MODEL 
+# 
+# 1. carve out a subset of the main data `pf_data`
+# ...
+# 
+# 2. get it into matrix format -- `input_matrix`
+# ...
+# 
+# 3. specify costs, priors, and alpha 
+# ...
+# 
+# 4. call `fullRecursion()` on the matrix, with params -- `output_matrix`
+# ... 
+# 
+# 5. visualize result 
+# make_rsa_plot(input_matrix, output_matrix)
 
-reason(input_matrix, 
-       costs=rep(0, ncol(input_matrix)), 
-       priors=rep(1, nrow(input_matrix)), 
-       alpha=1, 
-       depth=2, 
-       recycle_priors=FALSE)
+### STEPS FOR DOING MORE THAN ONE ITERATION 
+# prep the data to an `input_matrix`, 
+# specify parameters, 
+# and then call `reason()` on it 
+```
 
-rsa.fullRecursion(input_matrix, 
-                  costs=rep(0, ncol(input_matrix)), 
-                  priors=rep(1, nrow(input_matrix)),
-                  alpha=1)
-fullRecursion(input_matrix, 
-              costs=rep(0, ncol(input_matrix)), 
-              priors=rep(1, nrow(input_matrix)),
-              alpha=1)
+## 4. Play around with parameters + see how that changes results!
 
+```{r}
+# HINT: TRY SETTING THE COSTS VERY HIGH AND ALPHA VERY LOW 
+semantics_matrix <- matrix(
+  nrow=5, ncol=3, dimnames=list(paste0("row", 1:5), c("none","some","all")), 
+  data=c(1.0, .00, .00, 
+         .00, .00, .00, 
+         .25, .25, .25, 
+         .25, .00, .00, 
+         .00, .00,  1)
+)
+
+costs <- rep(1, ncol(semantics_matrix))
+priors <- rep(.5, nrow(semantics_matrix))
+alpha <- 1
+depth <- 1
+recycle_priors <- TRUE
+
+# same as `rsa.reason()` but source code is cleaner 
+pragmatics_matrix <- reason(
+  semantics_matrix, 
+  costs=costs, priors=priors, alpha=alpha, depth=depth, 
+  recycle_priors=recycle_priors
+)
+
+make_rsa_plot(semantics_matrix, pragmatics_matrix)
+
+# same as `rsa.fullRecursion()` but source code is cleaner 
+# also same as `reason` but with depth=1
+fullRecursion(
+  input_matrix, 
+  costs=costs, priors=priors, alpha=alpha
+)
 ```
@@
 - $S_n$ selects utterance $u$ from a set of alternative utterances 
-- an actual ("pragmatic") **listener** ???
-- FINISH
+- an actual ("pragmatic") **listener** reasons about the meaning based upon the utility of the actual utterance given the state of the world 
+- ...
 
@@
 
-#### equation 1 (GF2012)
+#### Equation 1 (GF2012)
 
@@
 
-#### equation 2 (GF2012)
+<br>
+
+#### Equation 2 (GF2012)
 
@@
 
-##### equation S1 (GF12 supplement)
+<br>
+
+#### Equation S1 (GF12 supplement)
 
@@
 
-#### equation S2 -- utility (GF12 supplement)
+<br>
+
+#### Equation S2 -- utility (GF12 supplement)
 
@@
 
-##### surprisal (GF12 supplement)
+<br>
+
+#### Surprisal (GF12 supplement)
 
@@
 
-##### utility and surprisal (GF12 supplement)
+<br>
+
+#### Utility and surprisal (GF12 supplement)
 
@@
 
-##### equation S3 -- literal listener model derivation (GF12, supplementary materials)
+<br>
+
+#### Equation S3 -- literal listener model derivation (GF12, supplementary materials)
 
@@
     0             & \text{otherwise }
-    \end{cases}
-$$
+    \end{cases}$$
 
-##### equation S4 -- listener likelihood (GF12, supplement)
+<br>
+
+#### Equation S4 -- listener likelihood (GF12, supplement)
 
@@
 ```{r eval=FALSE, include=FALSE}
+# THIS IS JUST TIM MISC NOTES TO SELF!
+
 ### rrrsa::rsa.fullRecursion(m, costs, priors, alpha) ------
@@
 ```
+
+<style>
+<!-- /* use this to get font import code: http://fonts.googleapis.com/css?family=FONT+NAME*/ -->
+<!-- /* use **latin**, not latin-extended */ -->
+@font-face {
+  font-family: 'Source Sans Pro';
+  font-style: normal;
+  font-weight: 400;
+  src: local('Source Sans Pro Regular'), local('SourceSansPro-Regular'), url(https://fonts.gstatic.com/s/sourcesanspro/v11/ODelI1aHBYDBqgeIAH2zlJbPFduIYtoLzwST68uhz_Y.woff2) format('woff2');
+  unicode-range: U+0000-00FF, U+0131, U+0152-0153, U+02C6, U+02DA, U+02DC, U+2000-206F, U+2074, U+20AC, U+2212, U+2215;
+}
+
+@font-face {
+  font-family: 'Roboto Mono';
+  font-style: normal;
+  font-weight: 400;
+  src: local('Roboto Mono'), local('RobotoMono-Regular'), url(https://fonts.gstatic.com/s/robotomono/v5/hMqPNLsu_dywMa4C_DEpY4gp9Q8gbYrhqGlRav_IXfk.woff2) format('woff2');
+  unicode-range: U+0000-00FF, U+0131, U+0152-0153, U+02C6, U+02DA, U+02DC, U+2000-206F, U+2074, U+20AC, U+2212, U+2215;
+}
+
+body {
+  padding: 10px;
+  font-size: 12pt;
+  font-family: 'Source Sans Pro', sans-serif;
+}
+
+h4 {
+  margin-top: 32px;
+  font-weight: bold;
+}
+
+blockquote {
+  font-size: 12pt;
+  font-family: 'Source Sans Pro', sans-serif;
+}
+
+code {
+  font-family: 'Roboto Mono', monospace;
+  font-size: 13px;
+}
+
+pre {
+  font-family: 'Roboto Mono', monospace;
+  font-size: 13px;
+}
+
+.table {
+  width: 75%;
+  align: center;
+}
+
+h3.subtitle {
+  font-size: 26px;
+  /*color: #aca39a;*/
+  font-style: normal !important;
+  font-weight: lighter !important;
+  border-bottom: none;
+  padding-top: 4px !important;
+}
+
+h4.author { 
+  font-size: 22px;
+  color: #aca39a;
+  font-weight: lighter !important;
+  padding-top: 0px !important;
+  margin-top: 2px !important;
+  margin-bottom: 12px !important;
+  padding-bottom: 12px !important;
+}
+
+h3 {
+  margin-top: 42px;
+}
+
+h2 {
+  margin-top: 48px;
+  border-bottom: 2px solid; 
+}
+
+</style>

### 2017-10-22 paper01_peloquin_frank2016/model_walkthru/paper01_peloquin_frank2016.rmd
fix a few things start playing around

@@
 
+# dnorm(), qnorm(), pnorm() 
+# hist(rnorm(1000, mean=100, sd=1))
+
+# priors <- normVec(rnorm(nrow(semantics_matrix), mean=.5, sd=.1))
 costs <- c(none=0, some=0, all=0)
-priors <- rnorm(nrow(semantics_matrix), mean=.5, sd=.1)
+priors <- rep(.2, 5)
 alpha <- 1
@@
 # need to make sure length matches nrow 
-priors <- rnorm(nrow(input_matrix), mean=.5, sd=.1)
+# priors <- rnorm(nrow(input_matrix), mean=.5, sd=.1) 
+priors <- rep(.2, nrow(input_matrix))
 
@@
 ### 3.3 Specific example: [WHATEVER SEEMS MOST INTERESING!]
+
+```{r}
+### THIS CHUNK IS JUST A SANDBOXXXE -----
+
+# carve out a subset
+NSA <- pf_data %>% 
+  filter(dataset=="pf5") %>%
+  filter(scale=="some_all") %>% 
+  select(-exp, -scale, -dataset) %>% 
+  select(stars, words, study1, study2, speaker.p) %>% 
+  arrange(stars, words)
+
+# this is the semantic representation (prob dist) of each scalar item 
+NSA_matrix <- NSA %>% 
+  # study1 or study2 or speaker.p -- still not sure which col! 
+  select(stars, words, value=study1) %>% 
+  tidyr::spread(words, value) %>% 
+  set_rownames(.$stars) %>%  # `stars` now encoded as rownames
+  select(none, some, little, most, all) %>% 
+  as.matrix()
+
+# need to make sure length matches ncol
+costs <- setNames(rep(0, ncol(NSA_matrix)), nm=colnames(NSA_matrix))
+
+# need to make sure length matches nrow 
+priors <- normVec(rep(1, nrow(NSA_matrix)))
+
+# parameter that fixes "how much the cost matters" 
+alpha <- 1
+
+# run an iteration
+NSA_pragmatics <- fullRecursion(NSA_matrix, costs, priors, alpha)
+
+# visualize result
+make_rsa_plot(NSA_matrix, NSA_pragmatics)
+```
+
+### 3.4 Steps for doing one full iteration
 

### 2017-10-22 paper01_peloquin_frank2016/model_walkthru/rsa_sandbox.rmd
fix a few things start playing around

---
title: "rrrsa imp"
output: html_document
---

```{r setup}
lefftpack::lazy_setup()
```

```{r normalize}
normalize <- function(vec){
  # all elements of `vec` must be positive or zero 
  stopifnot(min(vec, na.rm=TRUE) >= 0)
  # if `vec` is all zeros, just return it right back 
  if (length(unique(vec))==1){
    if (unique(vec)==0){
      return(vec)
    }
  }
  # normalize by dividing each element by the sum 
  return(vec / sum(vec, na.rm=TRUE))
}
```

```{r normalize_matrix}
normalize_matrix <- function(mat, rows_or_cols){
  if (!rows_or_cols %in% c("rows", "cols")){
    message("gotta normalize over rows or cols mayne!")
    return(NULL)
  }
  if (rows_or_cols=="rows"){
    return(t(apply(mat, MARGIN=1, normalize)))
  } 
  if (rows_or_cols=="cols"){
    return(apply(mat, MARGIN=2, normalize))
  }
}
```

```{r utility}
utility <- function(items, costs, alpha){
  # `items` is a vector of nums, each representing a word/item 
  # `costs` must have same length as `items`, `alpha` a positive scalar

  # apply `informativity()` to each item with specified costs and alpha. 
  # then collect them all and return with the same shape/order as `items`
  inform_unnorm <- mapply(informativity, prob=items, cost=costs, alpha=alpha)

  # normalization here is just dividing each number by the sum of them all 
  inform_norm <- normalize(inform_unnorm)

  return(inform_norm)
}

```

```{r informativity}
informativity <- function(prob, cost, alpha){
  # `prob` must be a probability, `alpha` a positive scalar 
  inform <- ifelse(prob == 0, 0, exp(-alpha * (-log(prob) - cost)))
  return(inform)
}

```

```{r full_recursion}
full_recursion <- function(mat, costs, priors, alpha){

  cost_mat <- matrix(
    nrow=nrow(mat), ncol=ncol(mat), 
    dimnames=dimnames(mat), 
    data=rep(as.numeric(costs), times=1, each=nrow(mat))
  )

  # the likelihood function is `utility()` 
  likelihood <- t(sapply(1:nrow(mat), function(row){
    utility(items=mat[row, ], costs=cost_mat[row, ], alpha=alpha)
  }))

  unnorm_posterior <- apply(likelihood, MARGIN=2, function(col) priors * col)

  posterior <- normalize_matrix(unnorm_posterior, rows_or_cols="cols")

  dimnames(posterior) <- list(rownames(mat), colnames(mat))

  return(posterior)
}

```

```{r rsa_convenience}
rsa_matrix_to_df <- function(rsa_matrix, response_colname){
  out <- as.data.frame(rsa_matrix) %>% 
    mutate(quantity = as.numeric(gsub("row", "", rownames(rsa_matrix)))) %>% 
    melt(id.vars="quantity", variable.name="word") %>% 
    mutate(word = as.character(word)) %>% 
    arrange(quantity, word) 
    # rename(semantics = "value")

  names(out)[names(out)=="value"] <- response_colname

  return(out)
}

make_rsa_plot <- function(input_matrix, output_matrix, facet_words=TRUE){

  input_df <- rsa_matrix_to_df(input_matrix, response_colname="semantics")
  output_df <- rsa_matrix_to_df(output_matrix, response_colname="pragmatics")

  input_output <- 
    full_join(input_df, output_df, by=c("quantity", "word")) %>% 
    melt(id.vars=c("quantity", "word")) %>% 
    rename(type = variable) %>% 
    mutate(type = as.character(type))

  input_output$type <- factor(input_output$type, 
                              levels=c("semantics","pragmatics"))

  if (facet_words){
    the_plot <- 
      ggplot(input_output, aes(x=quantity, y=value, color=type)) + 
      geom_point(size=rel(.5)) + 
      geom_line() + 
      scale_color_manual(values=c("blue", "orange")) + 
      facet_wrap(~word)

  } else {
    the_plot <- 
      ggplot(input_output, aes(x=quantity, y=value, color=word)) + 
      geom_line() + 
      facet_wrap(~type)
  }

  return(the_plot)
}

rsaplot <- function(smat, costs, priors, alpha, facet_words=TRUE){
  if (!all(colSums(smat)==1)){
    smat <- normalize_matrix(smat, rows_or_cols="cols")
  }
  if (sum(priors)!=1){
    priors <- normalize(priors)
  }
  pmat <- full_recursion(smat, costs, priors, alpha)
  return(make_rsa_plot(smat, pmat, facet_words=facet_words))
}

```

```{r semantics_matrix}
words <- c("none","none_all","some","some_notall","most","all","not_all")
costs <- setNames(object=c(1, 1, 1, 1, 1, 1, 1), nm=words)

points <- 1:5
priors <- normalize(rep(1, length(points)))

alpha <- 1

semantics_unnorm <- matrix(
  data=rep(NA, length(points)*length(words)), 
  nrow=length(points), ncol=length(words), 
  dimnames=list(paste0("row", points), words)
)
semantics_unnorm[, "none"]        <- c(1, 0, 0, 0, 0)
semantics_unnorm[, "none_all"]    <- c(1, 0, 0, 0, 1)
semantics_unnorm[, "some"]        <- c(0, 1, 1, 1, 1)
semantics_unnorm[, "some_notall"] <- c(0, 1, 1, 1, 0)
semantics_unnorm[, "most"]        <- c(0, 0, 1, 1, 1)
semantics_unnorm[, "all"]         <- c(0, 0, 0, 0, 1)
semantics_unnorm[, "not_all"]     <- c(1, 1, 1, 1, 0)
semantics_unnorm

rsaplot(semantics_unnorm, costs, priors, alpha)

s_a_na <- matrix(
  data=c(0,1,1,1,1,   0,0,0,0,1,    1,1,1,1,0), 
  nrow=5, ncol=3, dimnames=list(paste0("row", 1:5), 
                                c("some","all","not_all"))
)
rsaplot(s_a_na, costs=c(1,1,1), priors=rep(1, 5), alpha=1)

s_a <- matrix(
  data=c(0,1,1,1,1,   0,0,0,0,1), 
  nrow=5, ncol=2, dimnames=list(paste0("row", 1:5), 
                                c("some","all"))
)
rsaplot(s_a, costs=c(1,1), priors=rep(1, 5), alpha=1)

s_na <- matrix(
  data=c(0,1,1,1,1,   1,1,1,1,0), 
  nrow=5, ncol=2, dimnames=list(paste0("row", 1:5), 
                                c("some","not_all"))
)
rsaplot(s_na, costs=c(1,1), priors=rep(1, 5), alpha=1)

```

### 2017-10-22 paper01_peloquin_frank2016/model_walkthru/annotated-source-code/fullRecursion.r

@@ -1,17 +1,28 @@
-fullRecursion <- function(m, costs, priors, alpha=1){
+# semantics_matrix <- matrix(
+#   nrow=5, ncol=3, dimnames=list(paste0("row", 1:5), c("none","some","all")), 
+#   data=c(1.0, .00, .00, 
+#          .00, .00, .00, 
+#          .25, .25, .25, 
+#          .25, .00, .00, 
+#          .00, .00,  1)
+# )
+# costs <- c(none=0, some=0, all=0)
+# priors <- rep(.2, 5)
+# alpha <- 1
+# fullRecursion(semantics_matrix, costs, priors, alpha)
+
+fullRecursion <- function(m, costs, priors, alpha){
   
   cost_mat <- matrix(
     nrow=nrow(m), ncol=ncol(m), 
-    dimnames=list(paste0("row", 1:5), c("none","some","all")), 
-    data=rep(costs, times=1, each=nrow(m))
+    dimnames=dimnames(m), 
+    data=rep(as.numeric(costs), times=1, each=nrow(m))
   )
   
-  likelihood <- t(mapply(utility, 
-                         items=split(m, f=rownames(m)), 
-                         costs=split(cost_mat, row(cost_mat)), 
-                         alpha=alpha))
-  # identical(cost_mat, costsAsMatrix)
-  
+  # the likelihood function is `utility()` 
+  likelihood <- t(sapply(1:nrow(m), function(row){
+    utility(items=m[row, ], costs=cost_mat[row, ], alpha=alpha)
+  })) 
   
   unnorm_posterior <- apply(likelihood, MARGIN=2, function(col) priors * col)
   
@@ -22,6 +33,17 @@ fullRecursion <- function(m, costs, priors, alpha=1){
   return(posterior)
 }
 
+### THIS IS HOW IT'S DONE ORIGINALLY -- THE SAPPLY WAY IS CLEANER IMO THO 
+# likelihood <- t(mapply(
+#   utility, 
+#   items=split(m, f=rownames(m)), 
+#   costs=split(cost_mat, row(cost_mat)), 
+#   alpha=alpha
+# ))
 
-
-
+# cost_mat <- matrix(
+#   nrow=nrow(m), ncol=ncol(m), 
+#   dimnames=list(paste0("row", 1:5), c("none","some","all")), 
+#   data=rep(costs, times=1, each=nrow(m))
+# )
+# identical(cost_mat, costsAsMatrix)

### 2017-10-22 paper01_peloquin_frank2016/model_walkthru/annotated-source-code/informativity.r

@@ -1,4 +1,3 @@
-
 informativity <- function(m_u, cost, alpha){
   # `m_u` must be a probability, `alpha` a positive scalar 
   inform <- ifelse(m_u == 0, 0, exp(-alpha * (-log(m_u) - cost)))
@@ -11,7 +10,16 @@ informativity <- function(m_u, cost, alpha){
 # 
 # n <- 100
 # 
-# use combn() or outer() to put together
+# use combn() or outer() to put together??
 # then plot it by varying each param 
 
-
+# 
+# function (m_u, alpha = 1, cost = 0) 
+# {
+#   if (m_u < 0 | m_u > 1) 
+#     stop("Invalid semantic 'm_u' value, must be between [0, 1]")
+#   if (alpha < 0) 
+#     stop("Invalid alpha, must be a positive, non-zero numerical expression")
+#   ifelse(m_u == 0, 0, exp(-alpha * (-log(m_u) - cost)))
+# }
+# 

### 2017-10-22 paper01_peloquin_frank2016/model_walkthru/annotated-source-code/normVec.r

@@ -8,3 +8,7 @@ normVec <- function(v){
     return(v/sum(v, na.rm=TRUE))
   }
 }
+
+# vec <- 1:10
+# sum(normVec(vec))
+# plot(vec, normVec(vec))

### 2017-10-22 paper01_peloquin_frank2016/model_walkthru/annotated-source-code/utility.r

@@ -11,6 +11,10 @@ utility <- function(items, costs, alpha){
   return(inform_norm)
 }
 
+# sapply(list(1,2,3,4), function(x) x^2)
+# lapply(list(1,2,3,4), function(x) x^2)
+
+# ?mapply()
 
 
 # ### ORIGINAL 

### 2017-10-22 paper01_peloquin_frank2016/model_walkthru/setup.r

@@ -6,8 +6,8 @@ source("annotated-source-code/utility.r")
 source("annotated-source-code/informativity.r")
 source("annotated-source-code/fullRecursion.r")
 source("annotated-source-code/reason.r")  
-# source("annotated-source-code/runDf.r") # DOES NOT EXISST HYETS!
 source("annotated-source-code/tuneDepthAlpha.r")
+# source("annotated-source-code/runDf.r") # DOES NOT EXISST HYETS!
 
 
 ### CONTENTS OF `rrrsa::` PACKAGE ---------------------------------------------
@@ -15,12 +15,12 @@ source("annotated-source-code/tuneDepthAlpha.r")
 # FIVE DATASETS
 rsa_data <- list(
   # from frank et al submitted
-  frank = d_pragmods,
+  frank = rrrsa::d_pragmods,
   # from pelfrank16 
-  pf2   = peloquinFrank_2Alts,
-  pf3   = peloquinFrank_3Alts,
-  pf4   = peloquinFrank_4Alts,
-  pf5   = peloquinFrank_5Alts
+  pf2   = rrrsa::peloquinFrank_2Alts,
+  pf3   = rrrsa::peloquinFrank_3Alts,
+  pf4   = rrrsa::peloquinFrank_4Alts,
+  pf5   = rrrsa::peloquinFrank_5Alts
 )
 
 # THREE HELPER FUNCTIONS

### 2017-11-03 rsa-math.rmd
put up materials for LG15 and FG12 nov03

--- 2017-10-20 paper01_peloquin_frank2016/model_walkthru/paper01_peloquin_frank2016.rmd
@@
 ---
-title: "Model Walkthru for Peloquin & Frank (2016)"
-subtitle: "\"Determining the alternatives for scalar implicature\""
-author: "timothy leffel, oct20/2017"
-output: 
-  html_document:
-    css: style.css
-    self_contained: no
+title: ""
+output: pdf_document
+font_size: 10pt
+geometry: margin=.25in
 ---
 
-```{r setup}
-source("setup.r")
-```
+## RSA math cheatsheet (material from Frank and Goodman 2012 -- see paper for context)
 
-## Quick recap of paper
-
-## Contents of model implementation package
-
-Below is a listing of the contents of the `rrrsa::` package.
-
-#### Five datasets 
-```{r}
-# FIVE DATASETS
-rsa_data <- list(
-  # from frank et al submitted:
-  frank = rrrsa::d_pragmods,
-  # from pelfrank16: 
-  pf2   = rrrsa::peloquinFrank_2Alts,
-  pf3   = rrrsa::peloquinFrank_3Alts,
-  pf4   = rrrsa::peloquinFrank_4Alts,
-  pf5   = rrrsa::peloquinFrank_5Alts
-)
-```
-
-#### Three helper functions 
-```{r}
-# THREE HELPER FUNCTIONS
-rsa_helper <- list(
-  # converts vec2 to whatever type vec1 is
-  convertVecType = rrrsa::rsa.convertVecType, 
-  # this is just `function(x) x/sum(x)` for positive `x`
-  normVec        = rrrsa::rsa.normVec, 
-  # just changes desired column names 
-  renameCol      = rrrsa::rsa.renameCol
-)
-```
-
-#### Six analysis functions 
-```{r}
-# SIX ANALYSIS FUNCTIONS
-rsa_model <- list(
-  # computes informativity given params `m_u`, `alpha`, `cost`
-  informativity  = rrrsa::rsa.informativity,
-  # computes utility given params `items`, `costs`, `alpha`
-  utility        = rrrsa::rsa.utility,
-  # wrapper that runs `depth`-many iterations of `rsa.fullRecursion()` 
-  reason         = rrrsa::rsa.reason,
-  # wrapper around `rsa.reason` that accepts input as a data frame 
-  runDf          = rrrsa::rsa.runDf,
-  # explores correlation between data and model preds for varying `alpha` 
-  tuneDepthAlpha = rrrsa::rsa.tuneDepthAlpha,
-  # main function implementing RSA model 
-  fullRecursion  = rrrsa::rsa.fullRecursion
-)
-```
-
-## An example in action
-
-```{r}
-### ADAPTED CODE FROM PACKAGE VIGNETTE FOR DEMO OF fullRecursion()
-
-# m/dat_pre ~~> matrix of semantics (rows=meaning (m rows), cols=words (n cols))
-# costs     ~~> ncol(m) vector of costs (default is 0 valued vector)
-# priors    ~~> nrow(m) vector of priors (default is uniform)
-# alpha     ~~> decision hyper-param
-dat_pre <- matrix(
-  nrow=5, ncol=3, dimnames=list(paste0("row", 1:5), c("none","some","all")), 
-  data=c(1.0, .00, .00, 
-         .00, .00, .00, 
-         .25, .25, .25, 
-         .25, .00, .00, 
-         .00, .00,  1)
-)
-dat_pre_df <- rsa_matrix_to_df(dat_pre, response_colname="semantics")
-
-costs <- c(none=0, some=0, all=0)
-priors <- rnorm(nrow(dat_pre), mean=.5, sd=.1)
-alpha <- 1
-
-# rsa.fullRecursion(m=m, costs=rep(0, ncol(m)), priors=rep(1, nrow(m)), alpha=2)
-dat_post <- fullRecursion(m=dat_pre, 
-                          costs=rep(0,ncol(dat_pre)), 
-                          priors=rep(1,nrow(dat_pre)),
-                          alpha=alpha)
-
-knitr::kable(dat_pre, row.names=TRUE)
-knitr::kable(dat_post, row.names=TRUE)
-
-make_rsa_plot(dat_pre, dat_post)
-```
-
-## Reproducing the paper results
-
-#### get acquainted with the datasets, combine them
-
-```{r}
-source("setup.r")
-
-nams <- c("exp", "scale", "stars", "speaker.p", "words", "e11", "e6")
-#   1) quantityVarName :: "stars"
-#   2) semanticsVarName :: "speaker.p"
-#   3) itemVarName :: "words"
-#   4) groupVarName :: "scale"
-#   other) experiment number :: "exp"
-#   other) Pragmatic judgments study 1 (e6), study 2 (e11)
-sapply(rsa_data, dim)
-sapply(rsa_data, function(df){ identical(names(df), nams) })
-
-pf_data <- rsa_data[names(rsa_data)[names(rsa_data) != "frank"]]
-
-pf_data <- lapply(seq_along(pf_data), function(idx){
-  pf_data[[names(pf_data)[idx]]] %>% 
-    mutate(dataset = names(pf_data)[idx]) %>% 
-    mutate_if(is.factor, as.character) %>% 
-    rename(study1=e6, study2=e11)
-}) %>% 
-  (function(df_list) do.call(rbind, df_list))
-
-str(pf_data)
-# pf_data %>% filter(words=="none") %>% View() # *** 
-```
-
-#### Specific example: Expt1, three alts, <*none, some, all*>
-
-```{r}
-# carve out the <none, some, all> scale (both expts 1 and 2) 
-# pf_data %>% filter(scale=="some_all") %>% View()
-pf3_NSA <- pf_data %>% 
-  filter(dataset=="pf3") %>% 
-  filter(scale=="some_all") %>% 
-  select(-exp, -scale, -dataset) %>% 
-  select(stars, words, study1, study2, speaker.p) %>% 
-  arrange(stars, words)
-
-# all response fields are 0 for none, across the whole scale...
-# pf3_NSA %>% filter(words=="none") %>% View() # ***
-
-pf3_NSA_study1 <- pf3_NSA %>% 
-  select(stars, words, value=study1) %>% 
-  tidyr::spread(words, value) %>% 
-  set_rownames(.$stars) %>%  # `stars` now encoded as rownames
-  select(none, some, all) %>% 
-  as.matrix()
-
-# this is the semantic representation (prob dist) of each scalar item 
-input_matrix <- pf3_NSA_study1
-
-# need to make sure length matches ncol
-costs <- c(none=0, some=0, all=0)
-
-# need to make sure length matches nrow 
-priors <- rnorm(nrow(input_matrix), mean=.5, sd=.1)
-
-# parameter that fixes "how much the cost matters" 
-alpha <- 1
-
-# calculate the posterior distributions for each scalar item 
-output_matrix <- fullRecursion(
-  input_matrix, costs=costs, priors=priors, alpha=alpha
-)
-
-knitr::kable(input_matrix, row.names=TRUE)
-knitr::kable(output_matrix, row.names=TRUE)
-
-make_rsa_plot(input_matrix, output_matrix)
-```
-
-#### Specific example: Expt2, three alts, <*none, some, all*>
-
-#### Specific example: ...
-
-## Play around with parameters
-
-```{r}
-# play around with all the params!!! 
-rsa.reason(m=input_matrix, 
-           costs=rep(0, ncol(input_matrix)), 
-           priors=rep(1, nrow(input_matrix)), 
-           depth=2, 
-           alpha=1, 
-           usePriorEveryRecurse=FALSE)
-
-reason(input_matrix, 
-       costs=rep(0, ncol(input_matrix)), 
-       priors=rep(1, nrow(input_matrix)), 
-       alpha=1, 
-       depth=2, 
-       recycle_priors=FALSE)
-
-rsa.fullRecursion(input_matrix, 
-                  costs=rep(0, ncol(input_matrix)), 
-                  priors=rep(1, nrow(input_matrix)),
-                  alpha=1)
-fullRecursion(input_matrix, 
-              costs=rep(0, ncol(input_matrix)), 
-              priors=rep(1, nrow(input_matrix)),
-              alpha=1)
-
-```
-
-## Concepts used in RSA models
-
-#### primitive principles
-
-- speech-acts are rational behaviors -- actions in **signaling games** 
-- utterances are selected by their **expected utility** 
-- the utility of an utterance is a function of its **informativity** and its **cost**
-
-#### model of a conversational exchange
-
-- a **speaker** $S_n$ wants to communicate some proposition 
-- $S_n$ selects utterance $u$ from a set of alternative utterances 
-- an actual ("pragmatic") **listener** ???
-- FINISH
-
-## The mathematics underlying RSA models
-
-The following equations are from Goodman & Frank's (2012) model specification.
-
-#### equation 1 (GF2012)
+#### Equation 1 (FG2012) -- posterior probability
 
@@
 
-where
+*(i)* $r_s$ is an intended referent (element of $C$); *(ii)* $C$ is a context (set of objects $r, r', ...\in C$); *(iii)* $w$ is the word uttered by the speaker; *(iv)* the **prior** is $P(r_s)$; *(v)* the **likelihood** is $P(w|r_s, C)$; and *(vi)* the **normalizing constant** is $\sum\limits_{r'\in C} P(w|r', C) \times P(r')$
 
-- $r_s$ is an intended referent (element of $C$), 
-- $C$ is a context (set of objects $r, r', ...\in C$), and 
-- $w$ is the word uttered by the speaker. 
-- The **prior** is $P(r_s)$, 
-- the **likelihood** is $P(w|r_s, C)$, and 
-- the **normalizing constant** is $\sum\limits_{r'\in C} P(w|r', C) \times P(r')$
-
-#### equation 2 (GF2012)
+#### Equation 2 (FG2012) -- likelihood for FG12 (derived from formulas below)
 
@@
 
-where 
 - $|w|$ is the number of objects in $C$ to which word $w$ could(?) apply, 
@@
 
-##### equation S1 (GF12 supplement)
+#### Equation S1 (FG12 supplement)
 
@@
 
-where
+*(i)* the **context** $C = \{o_1, ..., o_n\}$ is a set of objects; *(ii)* the **vocabulary** is a set of words $V = \{w_1, ..., w_m\}$; *(iii)* each word $w\in V$ has a meaning $[\![ w ]\!] : C \rightarrow \{0, 1\}$; and *(iv)* words are chosen in proportion to their expected utility
 
-- the **context** $C = \{o_1, ..., o_n\}$ is a set of objects 
-- the **vocabulary** is a set of words $V = \{w_1, ..., w_m\}$ 
-- each word $w\in V$ has a meaning $[\![ w ]\!] : C \rightarrow \{0, 1\}$ 
-- words are chosen in proportion to their expected utility
-
-#### equation S2 -- utility (GF12 supplement)
+#### Equation S2 -- utility (FG12 supplement)
 
@@
 
-where
+where $I(w; r_s, C)$ is the informativeness of utterance $w$ w.r.t. intended referent $r_s \in C$; and $D(w)$ is $w$'s cost (in $C$?).
 
-- $I(w; r_s, C)$ is the informativeness of utterance $w$ w.r.t. intended referent $r_s \in C$, and 
-- $D(w)$ is $w$'s cost (in $C$?). (for single words, $D : V \rightarrow \mathbb{R}^+$ is constant)
+<br>
 
-##### surprisal (GF12 supplement)
-
-for a sample $x$ from a known distribution $p(x)$, the surprisal of $x$ is
+#### Surprisal (FG12 supplement), for a sample $x$ from a known distribution $p(x)$ 
 $$I_p(x) = -log(p(x))$$
 
-##### utility and surprisal (GF12 supplement)
+<br>
 
-utility decreases with surprisal:
+#### Utility and surprisal (FG12 supplement) -- utility decreases with surprisal
 
@@
 
-##### equation S3 -- literal listener model derivation (GF12, supplementary materials)
+<br>
+
+#### Equation S3 -- literal listener model derivation (FG12, supplementary materials)
 
@@
 
-$$\widetilde{w}_C(o) = 
-    \begin{cases}
-    \frac{1}{|w|} & \text{if } w(o) = true\\\\
-    0             & \text{otherwise }
-    \end{cases}
-$$
+$$\widetilde{w}_C(o) = \frac{1}{|w|} \textrm{ if } w(o) = true \textrm{, and otherwise } 0$$
 
-##### equation S4 -- listener likelihood (GF12, supplement)
+<!-- $$\widetilde{w}_C(o) =  -->
+<!--     \begin{cases} -->
+<!--     \frac{1}{|w|} & \text{if } w(o) = true\\\\ -->
+<!--     0             & \text{otherwise } -->
+<!--     \end{cases}$$ -->
+
+<br>
+
+#### Equation S4 -- listener likelihood (FG12, supplement)
 
@@
 
-"Thus in our experiments, the speaker's abstract goal of being informative reduces to a simple formulation: whoose a word that applies to the referent and picks out a relatively smaller section of the context.
-Listeners may then use this model of a speaker as their likelihood function, to be combined with prior information about contextual salience as in Equation 1 in the main text."
+<style>
+<!-- /* use this to get font import code: http://fonts.googleapis.com/css?family=FONT+NAME*/ -->
+<!-- /* use **latin**, not latin-extended */ -->
+@font-face {
+  font-family: 'Source Sans Pro';
+  font-style: normal;
+  font-weight: 400;
+  src: local('Source Sans Pro Regular'), local('SourceSansPro-Regular'), url(https://fonts.gstatic.com/s/sourcesanspro/v11/ODelI1aHBYDBqgeIAH2zlJbPFduIYtoLzwST68uhz_Y.woff2) format('woff2');
+  unicode-range: U+0000-00FF, U+0131, U+0152-0153, U+02C6, U+02DA, U+02DC, U+2000-206F, U+2074, U+20AC, U+2212, U+2215;
+}
 
-<!-- A `pragmatic listener` $P_{L_n}(m|u)$, reasons about intended meaning $m$ of an utterance $u$ by a `rational speaker` $P_{s_n}(u|m)$ who chooses an utterance according to the expected utility of an utterance $U(m;u)$.
-$\alpha$ is a decision noise parameter.
--->
+@font-face {
+  font-family: 'Roboto Mono';
+  font-style: normal;
+  font-weight: 400;
+  src: local('Roboto Mono'), local('RobotoMono-Regular'), url(https://fonts.gstatic.com/s/robotomono/v5/hMqPNLsu_dywMa4C_DEpY4gp9Q8gbYrhqGlRav_IXfk.woff2) format('woff2');
+  unicode-range: U+0000-00FF, U+0131, U+0152-0153, U+02C6, U+02DA, U+02DC, U+2000-206F, U+2074, U+20AC, U+2212, U+2215;
+}
 
-<!-- $$P_{L_n}(m|u) \propto P_{S_n}(u|m)P(m)$$ -->
-<!-- $$P_{S_n} \propto e^{U(m;u)}$$ -->
-<!-- $$U(m;u) = -\alpha(-\log(P_{L_{n-1}}(m|u)) - C(u))$$ -->
+body {
+  padding: 10px;
+  font-size: 12pt;
+  font-family: 'Source Sans Pro', sans-serif;
+}
 
-<!-- See Frank & Goodman (2012) and Goodman & Stuhmuller (2013) for the original descriptions of the RSA framework.
-Frank, et al.
-(Under Review) also provides a comprehensive presentation and evaluation of RSA.
--->
+h4 {
+  margin-top: 32px;
+  font-weight: bold;
+}
 
-```{r eval=FALSE, include=FALSE}
-### rrrsa::rsa.fullRecursion(m, costs, priors, alpha) ------
-# `m` a matrix
-# `costs` a vector of costs, w length == ncol(m) (defaults to zeros)
-# `priors` vector of priors, w length == nrow(m) (defaults to ones)
-# `alpha` scalar noise param, defaults to 1 
-# returns a posterior distribution 
-rrrsa::rsa.fullRecursion(m, costs, priors, alpha)
+blockquote {
+  font-size: 12pt;
+  font-family: 'Source Sans Pro', sans-serif;
+}
 
-# pdist=runif; nvals=100; ncols=4; params=c(0,1)
-fullrec <- function(pdist, nvals, ncols, params){
-  set.seed(3369)
-  m <- matrix(pdist(nvals, params[1], params[2]), ncol=ncols)
-  if (min(m) < 0) message("converting negative vals to positive")
-  m <- abs(m); m <- m/sum(m)
-  m_post <- rrrsa::rsa.fullRecursion(m=m)
-  par(mfrow=c(1,3)); print(hist(m)); print(hist(m_post)); print(plot(m, m_post))
-  title(paste0("with nvals=",nvals," ncols=",ncols, 
-               "\nparams=",paste(params, collapse=",")))
+code {
+  font-family: 'Roboto Mono', monospace;
+  font-size: 13px;
 }
-lapply(c(2,4,5,10,100), function(ncols){
-  fullrec(pdist=runif, nvals=1e3, ncols=ncols, params=c(0,1))
-})
-lapply(c(2,4,5,10,100), function(ncols){
-  fullrec(pdist=rnorm, nvals=1e3, ncols=ncols, params=c(0,1))
-})
 
-```
+pre {
+  font-family: 'Roboto Mono', monospace;
+  font-size: 13px;
+}
+
+.table {
+  width: 75%;
+  align: center;
+}
+
+h3.subtitle {
+  font-size: 26px;
+  /*color: #aca39a;*/
+  font-style: normal !important;
+  font-weight: lighter !important;
+  border-bottom: none;
+  padding-top: 4px !important;
+}
+
+h4.author { 
+  font-size: 22px;
+  color: #aca39a;
+  font-weight: lighter !important;
+  padding-top: 0px !important;
+  margin-top: 2px !important;
+  margin-bottom: 12px !important;
+  padding-bottom: 12px !important;
+}
+
+h3 {
+  margin-top: 42px;
+}
+
+h2 {
+  margin-top: 48px;
+  border-bottom: 2px solid; 
+}
+
+</style>

### 2017-11-03 rsa_sandbox.rmd
put up materials for LG15 and FG12 nov03

--- 2017-10-22 paper01_peloquin_frank2016/model_walkthru/rsa_sandbox.rmd
@@
 ---
-title: "rrrsa imp"
+title: "RSA sandbox"
+subtitle: "snippets and toy models (code here not tested extensively)"
+author: "timothy leffel // `tjleffel@gmail.com` // `http://lefft.xyz`"
+date: "updated nov03/2017"
 output: html_document
@@
 
+### setup
+
 ```{r setup}
 lefftpack::lazy_setup()
+# or use: 
+# library("magrittr"); library("dplyr"); library("reshape2"); library("ggplot2")
 ```
+
+### tweaks, doodles, etc. based on `rrrsa::` implementation
 
@@
   # the likelihood function is `utility()` 
+  # utility is just normalized informativity 
   likelihood <- t(sapply(1:nrow(mat), function(row){
@@
 ```
+
+### A little calculus refresher
+
+```{r}
+# dnorm(x, mean=0, sd=1): height of unit normal density at x
+plot(dnorm, from=-2, to=2)
+
+points <- seq(from=-2, to=2, by=.1)
+
+# proportion of "area under the curve" within the boundaries 
+integrate(dnorm, lower=-1.96, upper=1.96)$value 
+integrate(dnorm, -Inf, Inf)$value 
+integrate(dnorm, -Inf, 0)$value; integrate(dnorm, 0, Inf)$value
+
+# a little calculus refresher 
+dat <- data_frame(
+  point=points, 
+  norm_density=dnorm(points, mean=0, sd=1), 
+  integral_nd_nd=sapply(norm_density, function(x){
+    integrate(dnorm, lower=x-.05, upper=x+.05)$value
+  }), 
+  integral_nd_points=sapply(seq(from=-2, to=2, by=.1), function(x){
+    integrate(dnorm, lower=x-.05, upper=x+.05)$value
+  }), 
+  deriv_points=as.numeric(eval(deriv(~ dnorm(points), "points")))
+)
+dat %>% 
+  reshape2::melt(id.vars="point") %>% 
+  ggplot(aes(x=point, y=value, color=variable)) + 
+  geom_point() + facet_wrap(~variable, scales="free_y")
+
+# watch out -- since sum of points is (effectively) zero! 
+# sum(dat$point)
+lapply(dat, function(x) x/sum(x)) %>% lapply(sum)
+
+# plot(qnorm, from=-1, to=1)
+# plot(pnorm, from=-2, to=1)
+
+```
+
+```{r eval=FALSE, include=FALSE}
+### OLD CHUNK FROM PREVIOUS FIDDLE -- CONSOLIDATE AND/OR REMOVE AFTER REVIEWING
+
+### try out all funcs on some basic vals
+cost()                            # always maps to 1
+e()                               # euler's num
+eqn1(obj, utt, objects, words)    # .0952381
+eqn2(utt, obj, objects, words)    # .6666667
+
+s4_term(utt, objects)             # 1
+satisfies(obj, utt, objects)      # TRUE
+
+extension(utt, objects)           # "blue_circle"     
+extension("square", objects)      # "blue_square", "green_square"
+
+ll_interp(utt, obj, objects)     # TRUE
+prior_obj(obj, objects)          # .3333333
+prior_word(utt, words)           # .25
+
+words_applying_to_obj(words, obj, objects)               # "blue", "circle"
+
+prob_ref_given_word_in_context(obj, utt, objects, words) # .0952381
+prob_word_given_ref_in_context(utt, obj, objects, words) # .6666667
+
+sp_assertion(utt, obj, objects)  # word: "circle"; object: "blue_circle"; val=T
+
+# GOTTA FINALIZE DIST FUNC IN `informativeness()`, which this calls
+# *** utility(utt, obj, objects) ****************
+
+# THIS ALSO NEEDS FIXED (RELATED TO `utility()`)
+# *** informativeness(utt, dist_func=NA, obj)
+
+# THIS NEEDS FIXED IN THE FONXE FILE BEFORE CAN BE USED...
+# *** eqnS4(utt, obj, objects, words)   # 1.999856 <-- shd be same as eqn2()
+
+```
+
+<style>
+<!-- /* use this to get font import code: http://fonts.googleapis.com/css?family=FONT+NAME*/ -->
+<!-- /* use **latin**, not latin-extended */ -->
+@font-face {
+  font-family: 'Source Sans Pro';
+  font-style: normal;
+  font-weight: 400;
+  src: local('Source Sans Pro Regular'), local('SourceSansPro-Regular'), url(https://fonts.gstatic.com/s/sourcesanspro/v11/ODelI1aHBYDBqgeIAH2zlJbPFduIYtoLzwST68uhz_Y.woff2) format('woff2');
+  unicode-range: U+0000-00FF, U+0131, U+0152-0153, U+02C6, U+02DA, U+02DC, U+2000-206F, U+2074, U+20AC, U+2212, U+2215;
+}
+
+@font-face {
+  font-family: 'Roboto Mono';
+  font-style: normal;
+  font-weight: 400;
+  src: local('Roboto Mono'), local('RobotoMono-Regular'), url(https://fonts.gstatic.com/s/robotomono/v5/hMqPNLsu_dywMa4C_DEpY4gp9Q8gbYrhqGlRav_IXfk.woff2) format('woff2');
+  unicode-range: U+0000-00FF, U+0131, U+0152-0153, U+02C6, U+02DA, U+02DC, U+2000-206F, U+2074, U+20AC, U+2212, U+2215;
+}
+
+body {
+  padding: 10px;
+  font-size: 12pt;
+  font-family: 'Source Sans Pro', sans-serif;
+}
+
+h4 {
+  margin-top: 32px;
+  font-weight: bold;
+}
+
+blockquote {
+  font-size: 12pt;
+  font-family: 'Source Sans Pro', sans-serif;
+}
+
+code {
+  font-family: 'Roboto Mono', monospace;
+  font-size: 13px;
+}
+
+pre {
+  font-family: 'Roboto Mono', monospace;
+  font-size: 13px;
+}
+
+.table {
+  width: 75%;
+  align: center;
+}
+
+h3.subtitle {
+  font-size: 26px;
+  /*color: #aca39a;*/
+  font-style: normal !important;
+  font-weight: lighter !important;
+  border-bottom: none;
+  padding-top: 4px !important;
+}
+
+h4.author { 
+  font-size: 22px;
+  color: #aca39a;
+  font-weight: lighter !important;
+  padding-top: 0px !important;
+  margin-top: 2px !important;
+  margin-bottom: 12px !important;
+  padding-bottom: 12px !important;
+}
+
+h3 {
+  margin-top: 42px;
+}
+
+h2 {
+  margin-top: 48px;
+  border-bottom: 2px solid; 
+}
+
+</style>

### 2017-11-03 paper00_goodman_frank2012/FG12_funcs.r

@@ -1,0 +1,170 @@
+### FUNCTIONS FROM FRANK & GOODMAN (2012) SCIENCE PAPER 
+
+### posterior -----------------------------------------------------------------
+prob_ref_given_word_in_context <- function(object, word, objects, words){
+  message("calculate posterior density:\n", 
+          "  >> eqn1(",word,", ",object,", objects, words)\n")
+  
+  message("==========================================================\n",
+          "we will multiply the likelihood by the obj prior:\n", 
+          "    prob(", word, "|", obj, ") * prob(", obj, ") \n\n", 
+          "and divide by the sum of lhoods for ", 
+          word, " for all objects:\n",
+          "    sum o in objects: prob(", word, "|o)\n", 
+          "==========================================================\n")
+  eqn1(object=object, word, objects, words)
+}
+
+prob_word_given_ref_in_context <- function(word, object, objects, words){
+  message("calculate likelihood:\n", 
+          "  >> eqn2(", word, ", ", object, ", objects, words)")
+  eqn2(word, object, objects, words)
+}
+
+
+
+
+# eqn1(obj, word, objects, words) = 
+#      (eqn2 applied to the args) * prior(obj, objects) 
+#   ------------------------------------------------------
+#   sum for i in objects: [eqn2(word, i, objects, words)]
+eqn1 <- function(object, word, objects, words){
+  
+  # to get the posterior prob of object given word:
+  
+  # calculate the likelihood -- prob of word given object (equation 2)
+  likhood <- eqn2(word, object, objects, words) 
+  
+  # calculate the prior prob of object 
+  prior <- prior_obj(object, objects)
+  
+  # multiply the likelihood and prior 
+  numer <- likhood * prior 
+  
+  # calculate the likelihood for `word` for every object, and sum them up 
+  denom <- sum(
+    sapply(objects, function(r_prime){
+      eqn2(word, r_prime, objects, words)
+    })
+  )
+  
+  # normalize (likhood*prior) by the sum of the likelihoods 
+  out <- numer/denom
+  
+  # show the final calculation 
+  message(">> numer of eqn1:\n    (likhood * obj_prior) = ", numer)
+  message(">> denom of eqn1:\n    sum of the (likhood * obj_priors) = ", denom)
+  message(">> value of eqn1 for ", word, " and ", object, 
+          "\n  >> numer/denom = ", out, "\n")
+  
+  return(out)
+}
+
+
+
+
[…]

### 2017-11-03 paper00_goodman_frank2012/FG12_sandbox.r

@@ -1,0 +1,79 @@
+lefftpack::lazy_setup()
+source("FG12_funcs.r")
+
+# words that could be used
+words <- c("blue","green","fun",  "square","dog","owl","emu")
+# objects that could be referred to
+objects <- c("blue_square","blue_owl","blue_dog","blue_emu",
+             "green_square","green_emu", "green_dog", 
+             "fun_dog", "fun_emu", 
+             "square_owl")
+
+
+# get a random obj and a random utt
+# obj <- random_referent(objects)  # --> "square_owl"
+# utt <- random_utterance(words)   # --> "square"
+# LETS JUST DECIDE TO USE THIS EXAMPLE INSTEAD: 
+obj <- "square_owl"
+utt <- "square"
+
+sapply(words, function(w) length(extension(w, objects=objects)))
+
+
+c(`using object:` = obj, `   using utterance:` = utt)
+##################### 
+
+# ten objects, so object prior is 1/10 = .1  
+(sq_owl_prior <- prior_obj(obj, objects))
+
+# seven words, so utterance prior is 1/7 = .143 
+(sq_prior <- prior_word(utt, words))
+
+
+# probability of choosing 'square' to refer to the square owl 
+(likhood <- eqn2(utt, obj, objects, words))
+
+# the normalizing constant we use to get a probability for the posterior 
+(norm_constant <- sum(sapply(objects, function(o) eqn2(utt,o,objects,words))))
+
+# now we can calculate the posterior prob of referring to the square owl 
+# by saying 'square' in the context via bayes rule: 
+(sq_owl_posterior <- (sq_owl_prior * likhood) / norm_constant)
+
+
+# this is the same thing as just using equation 1: 
+eqn1(obj, utt, objects, words)
+
+# which is wrapped by 
+prob_ref_given_word_in_context(obj, utt, objects, words)  # .0714 
+
+square_owl_prior <- prior_obj(obj, objects)
+the_likhood      <- eqn2(utt, obj, objects, words) 
+norm_constant    <- sum(sapply(objects, function(o) eqn2(utt,o,objects,words)))
+square_owl_prior * the_likhood / norm_constant
+
+
+
+
+
+
+
+prob_word_given_ref_in_context("square", "square_owl", objects, words)
+# just applies equation 2: 
+#   >> say we have "square owl", call it SO 
+#   >> there are ten words, and eight objects 
[…]

### 2017-11-03 paper02_lassiter_goodman2015/LG15_funcs.r

@@ -1,0 +1,162 @@
+############ FUNCTIONS FOR LASSITER AND GOODMAN (2015) PAPER ------------------
+
+# number of cookies eaten, priors for each value 
+prior_A <- function(A){
+  stopifnot(A %in% 0:6)
+  if (A==0)
+    return(.94) else return(.01)
+}
+
+
+cost_u <- function(u){
+  stopifnot(u %in% c("none", "some", "all"))
+  return(2)
+  # technically it's this but let's keep it simple 
+  # 2/3 * length(u)
+}
+
+
+### 2.3 LITERAL LISTENER 
+# ex9, page 8 
+# The literal listener L0 is defined as an agent who responds to an utterance 
+# u in two steps: calculate  u , the literal interpretation of u in the 
+# relevant language, and condition the prior information state on 
+# the truth of  u .
+L0_prob <- function(A, u){
+  stopifnot(A %in% 0:6, u %in% c("none", "some", "all"))
+  # probability of state A given that [[u]] is true 
+  # e.g. probability that i ate 3 cookies given that 'some' is true 
+  if (u=="none"){
+    if (A==0) return(1) else return(0)
+  }
+  if (u=="some"){
+    if (A==0) return(0) else return(1/6)
+  }
+  if (u=="all"){
+    if (A==6) return(1) else return(0)
+  }
+}
+
+### 2.4 SPEAKER 
+S1_utility <- function(u, A){
+  stopifnot(A %in% 0:6, u %in% c("none", "some", "all"))
+  # utility of utterance u given state A 
+  # e.g. utility of saying 'some' given i ate 3 cookies 
+  l0 <- L0_prob(A, u)
+  if (l0==0){
+    return(0)
+  } else {
+    return(log(l0) - cost_u(u))
+  }
+}
+
+
+S1_prob <- function(u, A, uAlt, alpha=4, norm=FALSE){
+  # probability of choosing utterance u from Alt, given state A 
+  # e.g. probability of saying 'some' given i ate 3 
+  numerator <- exp(alpha * S1_utility(u, A))
+  
+  denominator <- sum(sapply(uAlt, function(u_prime){
+    exp(alpha * S1_utility(u_prime, A))
+  }))
+  
+  if (!norm){
+    return(numerator)
+  } else {
+    return(numerator / denominator)
+  }
+}
+
+
+### 2.5 PRAGMATIC LISTENER 
+L1_prob <- function(A, u, uAlt, A_space, alpha=4, norm=FALSE){
[…]

### 2017-11-03 paper02_lassiter_goodman2015/LG15_sandbox.r

@@ -1,0 +1,154 @@
+lefftpack::lazy_setup()
+source("LG15_funcs.r")
+
+
+### FIDDLING WITH LG15 EXAMPLES -----------------------------------------------
+
+### SCALAR IMPLICATURE EXAMPLE 
+u_space <- c("none", "some", "all") # the space of u's (words)
+A_space <- c(0, 1, 2, 3, 4, 5, 6) # the space of A's (num cookies)
+
+### p11
+# - A = 0 
+L0_prob(A=0, u="none") == 1
+L0_prob(A=0, u="some") == 0
+L0_prob(A=0, u="all") == 0
+
+S1_prob(u="some", A=1, uAlt=u_space)
+
+# PATTERN IS RIGHT BUT 0/1 FLIPPED! 
+sapply(u_space, function(u){
+  sapply(A_space, function(A){
+    S1_prob(u=u, A=A, uAlt=u_space, alpha=4)
+  })
+})
+
+# THE WRONG SIGN IN HERE THIS SEEMS TO BE THE CULPRIT(?!) 
+# (but also qualitatively seems weird...)
+sapply(u_space, function(u){
+  sapply(A_space, function(A){
+    S1_utility(u, A)
+  })
+})
+
+# SAME SIGN PROBLEM HERE 
+# probability of A given that u is true 
+sapply(u_space, function(u){
+  sapply(A_space, function(A){
+    try(L1_prob(A=A, u=u, uAlt=u_space, A_space=A_space, alpha=4, norm=TRUE))
+  })
+}) %>% set_rownames(paste0(0:6, "_cookies"))
+
+
+
+### REPRODUCE SOME EXAMPLES FROM THE PAPER 
+# u="none"; A=0
+eqn16 <- function(u, A){
+  if (L0_prob(A, u) == 0){
+    return(0)
+  } else {
+    exp(4 * (log(L0_prob(A, u)) - 4))
+  }
+}
+eqn16(u="all", A=0)
+eqn16(u="some", A=0)
+eqn16(u="none", A=0) # shd be 1.12e-7 ish 
+# 
+
+### [THIS BLOCK IS RIGHT BUT HAS A QUIRK]
+### p12, (18)
+L0_prob(A=1, u="some") # shd be 1/6 
+# 
+## THIS COMES OUT RIGHT (19), P13 [BUT SAYS IT IS NON-NORMALIZED...]
+L1_prob(A=1, u="some", uAlt=u_space, A_space=A_space, norm=TRUE) # .01 
[…]

### 2017-11-03 paper02_lassiter_goodman2015/LG15_funcs.r
added a bunch of notes during nov03 session, figgered a bunch of stuff out as a group, next thing is finish up LG15 basic stuff

@@ -8,9 +8,9 @@ prior_A <- function(A){
 }
 
 
-cost_u <- function(u){
+cost_u <- function(u, force_to_be=2){
   stopifnot(u %in% c("none", "some", "all"))
-  return(2)
+  return(force_to_be)
   # technically it's this but let's keep it simple 
   # 2/3 * length(u)
 }
@@ -54,10 +54,12 @@ S1_utility <- function(u, A){
 S1_prob <- function(u, A, uAlt, alpha=4, norm=FALSE){
   # probability of choosing utterance u from Alt, given state A 
   # e.g. probability of saying 'some' given i ate 3 
-  numerator <- exp(alpha * S1_utility(u, A))
+  numerator <- ifelse(
+    S1_utility(u, A) == 0, 0, exp(alpha * S1_utility(u, A))
+  )
   
   denominator <- sum(sapply(uAlt, function(u_prime){
-    exp(alpha * S1_utility(u_prime, A))
+    ifelse(S1_utility(u_prime, A)==0, 0, exp(alpha * S1_utility(u_prime, A)))
   }))
   
   if (!norm){
@@ -86,8 +88,9 @@ L1_prob <- function(A, u, uAlt, A_space, alpha=4, norm=FALSE){
   }
 }
 
+
 cost_u <- function(u){
-  return(2)
+  return(4)
 }
 
 
@@ -157,6 +160,38 @@ rsaplot <- function(smat, costs, priors, alpha, facet_words=TRUE){
 }
 
 
+# NOTE: THIS IS MODIFIED TO FIT EQNS 10 AND 11 FROM LG15 
+informativity <- function(prob, cost, alpha=1){
+  # `prob` must be a probability 
+  stopifnot(prob >= 0, prob <= 1)
+  inform <- ifelse(prob == 0, 0, exp(alpha * (log(prob) - cost)))
+  return(inform)
+}
+
+# alpha determines/affects how much the probability of an event/proposition/thingy affects the informativity
+probz <- seq(from=0, to=1, by=.01) 
+plot(probz, sapply(probz, function(p) informativity(prob=p, cost=1, alpha=.01)))
+
+plot(probz, sapply(probz, function(p) informativity(prob=p, cost=100, alpha=1)))
+# the more costly an utterance is, then the less the probability matters for its informativity 
+
+
+sapply(probz, function(p) informativity(prob=p, cost=1, alpha=1))
+
+plot(function(p) informativity(prob=p, cost=1, alpha=1), from=0, to=1)
+
[…]

### 2017-11-03 paper02_lassiter_goodman2015/LG15_sandbox.r

@@ -1,4 +1,4 @@
-lefftpack::lazy_setup()
+lefftpack::lazy_setup() # dplyr, magrittr, reshape2, ggplot2 
 source("LG15_funcs.r")
 
 
@@ -6,22 +6,58 @@ source("LG15_funcs.r")
 
 ### SCALAR IMPLICATURE EXAMPLE 
 u_space <- c("none", "some", "all") # the space of u's (words)
-A_space <- c(0, 1, 2, 3, 4, 5, 6) # the space of A's (num cookies)
+A_space <- c(0, 1, 2, 3, 4, 5, 6)   # the space of A's (num cookies)
 
+### LITERAL LISTENER INTERPRETATION WITH L0_prob()
 ### p11
 # - A = 0 
-L0_prob(A=0, u="none") == 1
-L0_prob(A=0, u="some") == 0
-L0_prob(A=0, u="all") == 0
+L0_prob(A=0, u="none")
+L0_prob(A=0, u="some")
+L0_prob(A=0, u="all")
 
-S1_prob(u="some", A=1, uAlt=u_space)
 
-# PATTERN IS RIGHT BUT 0/1 FLIPPED! 
+sapply(u_space, function(u){
+  sapply(A_space, function(A) L0_prob(A=A, u=u))
+})
+
+### PRIOR OVER NUMBER OF COOKIES EATEN WITH A_prior()
+sapply(A_space, function(A) prior_A(A=A))
+
+
+### APPLY EQUATION 16 TO GET SOMETHING PROPORTIONAL TO THE SPEAKER PROB OF u|A
+# eqn 10 and 11 combined (utilty plugged into 11)
+eqn16 <- function(u, A){
+  if (L0_prob(A, u) == 0){
+    return(0)
+  } else {
+    exp(4 * (log(L0_prob(A, u)) - 4))
+  }
+}
+
+# the unnormalized prob 
+eqn16(u="none", A=0) 
+S1_prob(u="none", A=0, uAlt=u_space, norm=FALSE)
+
+# top of page12 says this should be the normalized prob of "none" given A=0
+eqn16(u="none", A=0) / sum(sapply(u_space, function(u) eqn16(u=u, A=0)))
+S1_prob(u="none", A=0, uAlt=u_space, norm=TRUE)
+
+
+
+# ex18, close to bottom of p12 [correct]
+L0_prob(A=1, u="some") # .166667
[…]
